---
title: "AOR Aggregation"
format: html
editor: source
---

```{r}
# libraries
library(readxl)
library(tidyverse)
library(janitor)
library(lubridate)
library(sf)
library(arrow)
library(knitr)
library(kableExtra)
library(dbscan)
```

Loading in cleaned ICE encounters dataset

```{r}
ice <- read_csv('data/ice_encounters_cleaned.csv')
```




## Cluster the offices
Read in responsible sites data
```{r}
offices0 <- st_read("data/us_ice_office_points.shp")
```

```{r}
this_year <- as.integer(format(Sys.Date(), "%Y"))

offices <- offices0 |>
  rename(office = responsi_1) |> 
  rename(dep_country = departure_) |>
  rename(criminality = encounter_) |>
  rename(cit_country = citizenshi) |>
  rename(aor = responsibl) |> 
  mutate(deported = ifelse(dep_country == "Not applicable", 0, 1),
         age = this_year - birth_year)
```

```{r}
offices <- offices |> 
  mutate(
    cit_country = str_to_title(cit_country),
    region = case_when(
      # --- Canada + Mexico ---
      cit_country == "Canada" ~ "Canada",
      cit_country == "Mexico" ~ "Mexico",
  
      # --- Caribbean (including dependencies) ---
      cit_country %in% c(
        "Jamaica", "Haiti", "Dominican Republic", "Cuba",
        "Puerto Rico", "US Virgin Islands", "British Virgin Islands",
        "Cayman Islands", "Aruba", "Curacao", "Bonaire",
        "Guadeloupe", "Martinique", "Bermuda", "Montserrat",
        "Anguilla", "Saint Martin", "Sint Maarten",
        "Antigua and Barbuda", "Dominica", "Grenada",
        "Saint Lucia", "Saint Vincent and the Grenadines",
        "Trinidad and Tobago", "The Bahamas", "Barbados",
        "Saint Kitts and Nevis", "Turks and Caicos Islands"
      ) ~ "Caribbean",
  
      # --- Central America ---
      cit_country %in% c(
        "Belize", "Guatemala", "Honduras", "El Salvador",
        "Nicaragua", "Costa Rica", "Panama"
      ) ~ "Central America",
  
      # --- South America ---
      cit_country %in% c(
        "Argentina", "Bolivia", "Brazil", "Chile", "Colombia",
        "Ecuador", "Guyana", "Paraguay", "Peru",
        "Suriname", "Uruguay", "Venezuela"
      ) ~ "South America",
  
      # --- Africa ---
      cit_country %in% c(
        "Algeria", "Angola", "Benin", "Botswana", "Burkina Faso",
        "Burundi", "Cabo Verde", "Cameroon", "Central African Republic",
        "Chad", "Comoros", "Congo", "Democratic Republic of the Congo",
        "Djibouti", "Egypt", "Equatorial Guinea", "Eritrea",
        "Eswatini", "Ethiopia", "Gabon", "Gambia", "Ghana",
        "Guinea", "Guinea-Bissau", "Ivory Coast", "Kenya",
        "Lesotho", "Liberia", "Libya", "Madagascar", "Malawi",
        "Mali", "Mauritania", "Mauritius", "Morocco", "Mozambique",
        "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe",
        "Senegal", "Seychelles", "Sierra Leone", "Somalia",
        "South Africa", "South Sudan", "Sudan", "Tanzania",
        "Togo", "Tunisia", "Uganda", "Zambia", "Zimbabwe"
      ) ~ "Africa",
  
      # --- Asia ---
      cit_country %in% c(
        # East Asia
        "China", "Japan", "South Korea", "North Korea", "Mongolia", "Taiwan",
        # Southeast Asia
        "Indonesia", "Thailand", "Vietnam", "Malaysia",
        "Singapore", "Philippines", "Cambodia", "Laos",
        "Myanmar", "Brunei", "Timor-Leste",
        # South Asia
        "India", "Pakistan", "Bangladesh", "Nepal",
        "Sri Lanka", "Bhutan", "Maldives",
        # Central Asia (your rule)
        "Kazakhstan", "Kyrgyzstan", "Tajikistan",
        "Turkmenistan", "Uzbekistan",
        # Middle East (your rule)
        "Saudi Arabia", "United Arab Emirates", "Qatar", "Kuwait",
        "Bahrain", "Oman", "Yemen", "Jordan",
        "Lebanon", "Syria", "Iraq", "Iran",
        "Israel", "Palestine", "Turkey"
       
      ) ~ "Asia",
  
      # --- Europe (includes Russia) ---
      cit_country %in% c(
        "United Kingdom", "Ireland", "France", "Germany", "Spain", "Portugal",
        "Italy", "Belgium", "Netherlands", "Luxembourg",
        "Switzerland", "Austria", "Poland", "Czech Republic",
        "Slovakia", "Hungary", "Slovenia", "Croatia",
        "Serbia", "Bosnia and Herzegovina", "Montenegro",
        "North Macedonia", "Albania", "Greece",
        "Bulgaria", "Romania", "Moldova", "Ukraine", "Belarus",
        "Lithuania", "Latvia", "Estonia",
        "Norway", "Sweden", "Finland", "Denmark", "Iceland",
        "Georgia", "Armenia", "Azerbaijan", "Russia"
      ) ~ "Europe",
  
      # --- Oceania ---
      cit_country %in% c(
        "Australia", "New Zealand", "Papua New Guinea",
        "Fiji", "Samoa", "Tonga", "Vanuatu", "Solomon Islands",
        "Kiribati", "Micronesia", "Palau",
        "Marshall Islands", "Nauru", "Tuvalu"
      ) ~ "Oceania",
  
      TRUE ~ NA_character_
    )
  )

```

```{r}
offices_clean <- offices |>
  drop_na(criminality, gender, age, region)

nrow(offices_clean)
```

### Clustering 
```{r}
num_encounters <- offices_clean |>
  group_by(office) |>
  summarize(num_encounters = n())

coords <- st_coordinates(num_encounters)

set.seed(42069)
# density based clustering
db_clusters <- dbscan(coords, eps=0.5, minPts=1)

db_clusters
```

```{r}
# add cluster ID to data
num_encounters$cluster <- as.factor(db_clusters$cluster)
```

```{r}
cluster_summary <- num_encounters |>
  group_by(cluster) |>
  summarize(agg_num_encounters = sum(num_encounters)) |>
  arrange((agg_num_encounters))
```

Join small clusters (<1000) to the nearest larger cluster
```{r}
small_clusters <- cluster_summary |>
  filter(agg_num_encounters < 1000)

large_clusters <- cluster_summary |>
  filter(agg_num_encounters >= 1000)

# computer nearest large cluster for each small cluster
nearest <- small_clusters |>
  rowwise() |>
  mutate(nearest_cluster = large_clusters$cluster[which.min(st_distance(geometry, large_clusters$geometry))])
```

Merge cluster assignments to office-level data
```{r}
merge_lookup <- nearest |>
  select(cluster, nearest_cluster)

num_encounters <- num_encounters |>
  st_drop_geometry() |> 
  left_join(merge_lookup, by="cluster") |>
  mutate(cluster_final = ifelse(!is.na(nearest_cluster), nearest_cluster, cluster)) |>
  select(-nearest_cluster, -geometry)


```

```{r}
num_encounters |>
  group_by(cluster_final) |>
  summarize(total_cluster_encounters = sum(num_encounters))
```

Merge with individual-level data
```{r}
offices_clustered <- offices_clean |>
  left_join(num_encounters, by = "office") |>
  select(-cluster) |>
  mutate(cluster_final = as.factor(cluster_final))
```


Make it spatial
```{r}
clusters_sf <- st_as_sf(offices_clustered , 
                     coords = c("office_lon", "office_lat"), 
                     crs = 4326) 
```

```{r}
st_crs(clusters_sf)
```

Export data with clustered offices for mapping
```{r}
st_write(clusters_sf, "data/clustered_offices.shp")

```

## Create county shapefile with atttributes

```{r}
aors_counties <- st_read('data/aors_counties/aors_counties.shp')
```

Loading in ACS data that has info on native/foreign born populations by county

```{r}
nat_cit <- read_csv('data/nat_cit_cleaned.csv')
#colnames(nat_cit)
```

Loading in Decennical Census data that has info on race and ethnicity by county

```{r}
race_ethn <- read_csv('data/race_ethn_cleaned.csv')
#colnames(race_ethn)
```

Merge with county shapefile

```{r}

county_info_merged <- aors_counties |>         
  left_join(nat_cit,  by="fips") |>
  left_join(race_ethn, by="fips")

```

```{r}

county_info_merged <- county_info_merged |>
  select("name", 
         "state_name",
         "state_abbr",
         "state_fips",
         "county_fip",
         "fips",
         "population",
         "pop_sqmi",
         "shape_leng",
         "shape_area",
         "aor",
         "prop_not_a_citizen",
         "prop_black",
         "prop_asian",
         "prop_hispanic_or_latino",
         "geometry")
```

```{r}
county_info_merged |> st_geometry() |> head()
```

```{r}
st_crs(aors_counties)
summary(st_coordinates(aors_counties))
summary(st_coordinates(county_info_merged))

```



```{r}
# save county info shp
st_write(county_info_merged, "data/county_info.shp")
```


## Merge county demographics with ICE data
```{r}
counties_offices_joined <- st_read('data/counties_offices_joined.shp')
```
Group by office, then take the means of the demographic variables and create a variable of the population density
```{r}
colnames(counties_offices_joined)
```
Some renaming so I don't go insane
```{r}
counties_offices_joined <- counties_offices_joined |>
  rename("p_black" = "prp_blc",
         "p_asian" = "prop_sn",
         "p_hisp" = "prp_h__",
         "p_noncit" = "prp_n__",
         "st_abbr" = "stt_bbr") |>
  select(-c("nearest_x", "nearest_y", "feature_x", "feature_y")) 
  
```

Group by office, then take the means of the demographic variables and create a variable of the population density
```{r}
office_county_agg <- counties_offices_joined |>
  group_by(office) |>
  reframe(avg_p_blk = mean(p_black),
         avg_p_asn = mean(p_asian),
         avg_p_his = mean(p_hisp),
         avg_p_nc = mean(p_noncit),
         avg_pop_sqm = mean(pop_sqm)) |>
  st_drop_geometry()
```

### Read in clusters point data to join on
```{r}
clustered_offices <- st_read("data/clustered_offices_pr.shp")
```

Join by office
```{r}
cnty_clus_offices <- clustered_offices |>
  left_join(office_county_agg, by = "office") 

  
```

Export
```{r}
st_write(cnty_clus_offices, "data/cnty_clus_offices.shp")
```


## This was used for descriptive stats: not part of modeling process
## Working with just the Encounters data: Create a dataset that is aggregated by AOR

```{r}
colnames(ice)
```

```{r}
# function to get statistical mode
get_mode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

```{r}
aor_agg <- ice |>
  group_by(aor) |>
  summarize(
    num_encounters = n(),
    num_responsible_sites = n_distinct(responsible_site),
    num_event_landmarks = n_distinct(event_landmark),
    pct_no_charges_or_convictions = mean(encounter_criminality == "3 Other Immigration Violator", na.rm = TRUE),
    mode_cit_country = get_mode(citizenship_country),
    pct_deported = mean(!(departure_country == "Not applicable"), na.rm = TRUE)
  )

```

```{r}
head(aor_agg)
```

### Join with AOR shapefile for mapping

```{r}
aor_shapes <- st_read("data/aor_dissolved_shapes/aor_shapefile_dissolved.shp")
```

I had to clean up this dissolved shapefile

```{r}
#aor_shapes <- aor_shapes |>
  #select(aor, shape_leng, shape_area, geometry)

```

```{r}
#st_write(aor_shapes, "data/aor_only_shapes/aor_shapes.shp")
```

```{r}
aor_merged <- aor_shapes |>
  right_join(aor_agg, by = "aor")
```

```{r}
head(aor_merged)
```


## Aggregate this to the AOR level

```{r}
aor_demog <- county_info_merged |>
  group_by(aor) |>
  summarize(prop_noncitizen = sum(estimate_not_a_citizen)/sum(population),
         tot_noncitizen = sum(estimate_not_a_citizen),
         tot_pop = sum(population),
         prop_hispanic_or_latino = sum(total_hispanic_or_latino)/sum(population),
         prop_black_any = sum(total_black_any)/sum(population),
         prop_asian_any = sum(total_asian_any)/sum(population)) |>
  st_drop_geometry()


colnames(aor_demog)
```

### Merge with Encounters dataset

```{r}
encounters_demog_merged <- aor_merged |>
  select(-c(shape_leng, shape_area)) |>
  left_join(aor_demog, by="aor")
```

```{r}
st_write(encounters_demog_merged, "data/encounters_demog_merged.shp")
```

```{r}
aor_descrip_stats <- encounters_demog_merged |>
  st_drop_geometry() |>
  kable() |>
  kable_styling("striped")

aor_descrip_stats
save_kable(aor_descrip_stats, file = "aor_descrip_stats.png")
```

```{r}
write_csv(aor_descrip_stats, "data/aor_descriptive_stats.csv")
```

## Working with just the Encounters data: Create a dataset that is aggregated by docket office (responsible site)

```{r}
resp_site_agg <- ice |>
  group_by(responsible_site) |>
  summarize(
    aor = unique(aor),
    num_encounters = n(),
    num_event_landmarks = n_distinct(event_landmark),
    pct_no_charges_or_convictions = mean(encounter_criminality == "3 Other Immigration Violator", na.rm = TRUE),
    mode_cit_country = get_mode(citizenship_country),
    pct_deported = mean(!(departure_country == "Not applicable"), na.rm = TRUE)
  )
```

```{r}
head(resp_site_agg)
```

