---
title: "Modeling Framework"
format: html
editor: source
---
```{r}
library(tidyverse)
library(sf)
library(VGAM)
library(lubridate)
library(lme4)
library(spdep)
library(dbscan)
library(car)
library(corrplot)
library(GWmodel)
library(raster)
library(paletteer)
library(caret)
library(pROC)
#library(glmmTMB)
library(viridis)
library(emmeans)
select <- dplyr::select
```

Read in clustered data
```{r}
cnty_clus_offices <- st_read("data/cnty_clus_offices.gpkg")

# some preprocessing
cnty_clus_offices <- cnty_clus_offices |>
  drop_na(criminality, gender, age, region, avg_p_blk, avg_p_asn , avg_p_his, avg_p_nc, avg_pop_sqm)

cnty_clus_offices$deported <- factor(cnty_clus_offices$deported, levels = c(0,1))

cnty_clus_offices$region <- factor(cnty_clus_offices$region)

cnty_clus_offices$cluster <- factor(cnty_clus_offices$cluster)

cnty_clus_offices$gender <- factor(cnty_clus_offices$gender)

cnty_clus_offices$criminality <- factor(cnty_clus_offices$criminality)

cnty_clus_offices$criminality <- relevel(cnty_clus_offices$criminality, ref = "3 Other Immigration Violator")

cnty_clus_offices$cluster <- relevel(cnty_clus_offices$cluster, ref = "43")

cnty_clus_offices <- cnty_clus_offices |>
  mutate(program_type = factor(case_when(final_program == "ERO Criminal Alien Program" ~ "ERO Criminal Alien Program",
                                     final_program == "287G Program" | final_program == "287g Task Force" ~ "287(g)",
                                     .default = "Other")))
cnty_clus_offices$program_type <- relevel(cnty_clus_offices$program_type, ref = "Other")

latin_am_regions = c("Central America", "Caribbean", "Mexico", "South America")

cnty_clus_offices <- cnty_clus_offices |>
  mutate(latin_cit = case_when(region == "Mexico" ~ 1,
                                  region == "South America" ~ 1,
                                  region == "Central America" | region == "Caribbean" ~ 1,
                                  !(region %in% latin_am_regions) ~ 0)) 
```

```{r}

table(cnty_clus_offices$program_summary)

```


### Check for multicollinearity
```{r}
numeric_vars <- cnty_clus_offices |>
  group_by(office) |>
  select(c(avg_p_blk, avg_p_asn , avg_p_his, avg_p_nc, avg_pop_sqm)) |>
  drop_na() |>
  st_drop_geometry()
```

None of correlations are above 0.8, so I can keep all variables in the model
```{r}
corr_mat <- cor(numeric_vars)
corrplot(corr_mat, type="upper", method="ellipse")
```
```{r}
corr_mat
```

## Basic logistic regression
```{r}
freq_table <- table(cnty_clus_offices$prog_287g)
prop.table(freq_table)
```

```{r}
#cnty_clus_offices$criminality <- factor(cnty_clus_offices$criminality)

log_reg <-glm(deported ~ criminality + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_p_nc + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(log_reg , file="models/log_reg.rds")
```

```{r}
log_reg <- readRDS("models/log_reg.rds")
```

```{r}
summary(log_reg)
```


### Moran's I on basic logistic regression
```{r}
cnty_clus_offices$log_resids <- log_reg$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. We conclude that there is clustering behavior of the residuals. 
```{r}
offices_grouped <- cnty_clus_offices |>
  group_by(office) |>
  mutate(avg_log_resid = mean(log_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_log_resid, mat2listw(w1, style="W"))
```



```{r}
table(cnty_clus_offices$prog_287g)
```


## Interacting criminality with cluster 
Cluster reference group: 167
```{r}

cluster_reg_crim <- glm(deported ~ criminality + cluster + criminality:cluster+ gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his +  avg_p_nc  + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_reg_crim, file="cluster_reg_crim.rds")
```


```{r}
cluster_reg_crim <- readRDS("models/cluster_reg_crim.rds")
```


```{r}
summary(cluster_reg_crim)
```

#### Get coefficients for the interaction term for convicted criminals
```{r}
#### Get coefficients for criminality effects
coefs <- summary(cluster_reg_crim)$coefficients
```


```{r}
# Extract main effect of convicted criminal (same for all clusters)
clstr_fixed_crim <- coefs["criminality1 Convicted Criminal", "Estimate"]

# Extract interaction terms (vary by cluster)
clstr_int_crim <- coefs[grep("^criminality1 Convicted Criminal:cluster", rownames(coefs)), , drop = FALSE]

# Set non-significant interactions to 0 (or NA if you prefer)
clstr_int_crim[clstr_int_crim[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0


# Create interaction dataframe
interaction_df <- data.frame(
  cluster = sub("criminality1 Convicted Criminal:cluster", "", rownames(clstr_int_crim)),
  clstr_crim_int = clstr_int_crim[, "Estimate"]
)

# Add the fixed effect (same for all clusters)
interaction_df <- interaction_df |>
  mutate(clstr_crim_fixed = clstr_fixed_crim)

# For clusters with no interaction term (reference cluster or not in model)
# Add them with interaction = 0
cluster_office_df <- cnty_clus_offices

all_clusters <- unique(cluster_office_df$cluster)  # Get all cluster IDs from your data
interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(interaction_df, by = "cluster") |>
  mutate(
    clstr_crim_fixed = clstr_fixed_crim,  # Same for all
    clstr_crim_int = replace_na(clstr_crim_int, 0),  # 0 if no interaction
    clstr_crim_int_sig = ifelse(clstr_crim_int == 0, 1, 0)
  )

# Calculate cluster-specific effect
interaction_df <- interaction_df |>
  mutate(cluster_crim_effect = clstr_crim_fixed + clstr_crim_int)
```



```{r}
cluster_office_df <- cnty_clus_offices |>
  group_by(office) |>
  select(c(office, cluster, geom, aor)) |>
  slice(1) |>
  ungroup()
```


```{r}
cluster_office_df <- cluster_office_df|>
  left_join(interaction_df, by = "cluster")
```


Most regions are signiicantly different from each other
```{r}
region_means <- emmeans(cluster_reg_crim, ~ region)
pairs(region_means, adjust = "tukey") # Or "bonferroni", "holm", etc.
```

### Moran's I on cluster:criminality model
```{r}
cnty_clus_offices$clstr_crim_resids <- cluster_reg_crim$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- cnty_clus_offices |>
  group_by(office) |>
  mutate(avg_clstr_crim_resid = mean(clstr_crim_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_clstr_crim_resid, mat2listw(w1, style="W"))
```



## Interacting criminality with aor
s### exploration: calculate the proportion noncriminal per aor
```{r}
cnty_clus_offices |>
  group_by(aor)|>
  summarize(prop_noncrim = mean(criminality == "3 Other Immigration Violator")) |>
  arrange(desc(prop_noncrim))
```


```{r}

aor_reg <- glm(deported ~ criminality + aor + criminality:aor + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_p_nc+  avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)


saveRDS(aor_reg, file="aor_reg.rds")
```

```{r}
aor_reg <- readRDS("models/aor_reg.rds")
```


```{r}
summary(aor_reg)
```


#### Get coefficients for the interaction term for criminals
```{r}
#### Get coefficients for criminality effects at AOR level
coefs <- summary(aor_reg)$coefficients

# Extract main effect of convicted criminal (same for all AORs)
aor_fixed_crim <- coefs["criminality1 Convicted Criminal", "Estimate"]

# Extract interaction terms (vary by AOR)
aor_int_crim <- coefs[grep("^criminality1 Convicted Criminal:aor", rownames(coefs)), , drop = FALSE]

# Set non-significant interactions to 0 (or NA if you prefer)
aor_int_crim[aor_int_crim[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0

# Create interaction dataframe
aor_interaction_df <- data.frame(
  aor = sub("criminality1 Convicted Criminal:aor", "", rownames(aor_int_crim)),
  aor_crim_int = aor_int_crim[, "Estimate"]
)

# Get all unique AORs from your data
all_aors <- unique(cluster_office_df$aor)

# Add the fixed effect and handle missing interactions
aor_interaction_df <- data.frame(aor = all_aors) |>
  left_join(aor_interaction_df, by = "aor") |>
  mutate(
    aor_crim_fixed = aor_fixed_crim,  # Same for all AORs
    aor_crim_int = replace_na(aor_crim_int, 0),  # 0 if no interaction (reference AOR)
    aor_crim_int_sig = ifelse(aor_crim_int == 0, 1, 0)
  )

# Calculate marginal effect
aor_interaction_df <- aor_interaction_df |>
  mutate(aor_crim_effect = aor_crim_fixed + aor_crim_int)


cluster_office_df <- cluster_office_df |>
  left_join(aor_interaction_df, by = "aor")
```

```{r}
aor_interaction_df_pct <- aor_interaction_df |>
  mutate(aor_crim_marg_pctdiff = exp(aor_crim_marg) * 100) 
```




### Moran's I on aor:criminality model
```{r}
cnty_clus_offices$aor_crim_resids <- aor_reg$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- cnty_clus_offices |>
  group_by(office) |>
  mutate(avg_aor_crim_resid = mean(aor_crim_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_aor_crim_resid, mat2listw(w1, style="W"))
```

## Interaction Proportion Hispanic:cluster
```{r}
cluster_hisp <- glm(deported ~ criminality +cluster + avg_p_his:cluster+ gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his +  avg_p_nc  + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_hisp, file="cluster_hisp.rds")
```


```{r}
summary(cluster_hisp)
```


```{r}
cluster_hisp <- readRDS("models/cluster_hisp.rds")
```



#### Get coefficients for the interaction term for prop hispanic
```{r}
#### Get coefficients for age effects at cluster level
coefs <- summary(cluster_hisp)$coefficients

# Extract main effect of age
cluster_fixed_hisp <- coefs["avg_p_his", "Estimate"]

# Extract interaction terms
cluster_int_hisp <- coefs[grep(":avg_p_his$", rownames(coefs)), , drop = FALSE]

# Set non-significant interactions to 0
cluster_int_hisp[cluster_int_hisp[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0


# Create interaction dataframe - strip cluster prefix and :age suffix
hisp_interaction_df <- data.frame(
  cluster = sub("cluster", "", sub(":avg_p_his$", "", rownames(cluster_int_hisp))),
  cluster_int_hisp = cluster_int_hisp[, "Estimate"]
)

# Get all unique clusters
all_clusters <- unique(cluster_office_df$cluster)

# Add fixed effect and handle missing interactions
hisp_interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(hisp_interaction_df, by = "cluster") |>
  mutate(
    clstr_hisp_fixed = cluster_fixed_hisp,
    cluster_hisp_int = replace_na(cluster_int_hisp, 0),
    clstr_hisp_int_sig = ifelse(clstr_hisp_int == 0, 1, 0)
  )

# Calculate marginal effect
hisp_interaction_df <- hisp_interaction_df |>
  mutate(cluster_hisp_effect = clstr_hisp_fixed + cluster_hisp_int)


# Join to main dataframe
cluster_office_df <- cluster_office_df |>
  left_join(hisp_interaction_df, by = "cluster")
```



## Interaction Proportion Asian:cluster
```{r}
cluster_asian <- glm(deported ~ criminality + cluster + avg_p_asn :cluster+ gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his +  avg_p_nc  + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_asian, file="cluster_asian.rds")
```

```{r}
cluster_asian <- readRDS("models/cluster_asian.rds")
```

```{r}
summary(cluster_asian)
```


#### Get coefficients for the interaction term for prop asian
```{r}
#### Get coefficients for age effects at cluster level
coefs <- summary(cluster_asian)$coefficients

# Extract main effect of age
cluster_fixed_asi <- coefs["avg_p_asn", "Estimate"]

# Extract interaction terms
cluster_int_asi <- coefs[grep(": avg_p_asn $", rownames(coefs)), , drop = FALSE]

# Set non-significant interactions to 0
cluster_int_asi[cluster_int_asi[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0


# Create interaction dataframe - strip cluster prefix and :age suffix
asi_interaction_df <- data.frame(
  cluster = sub("cluster", "", sub(": avg_p_asn $", "", rownames(cluster_int_asi))),
  cluster_int_asi = cluster_int_asi[, "Estimate"]
)


# Add fixed effect and handle missing interactions
asi_interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(asi_interaction_df, by = "cluster") |>
  mutate(
    clstr_asi_fixed = cluster_fixed_asi,
    cluster_asi_int = replace_na(cluster_int_asi, 0),
    clstr_asi_int_sig = ifelse(clstr_asi_int == 0, 1, 0)
  )

# Calculate marginal effect
asi_interaction_df <- asi_interaction_df |>
  mutate(cluster_asi_effect = clstr_asi_fixed + cluster_asi_int)


# Join to main dataframe
cluster_office_df <- cluster_office_df |>
  left_join(asi_interaction_df, by = "cluster")
```


## Interaction LatinAmerica:cluster

```{r}
freq_table <- table(cnty_clus_offices$latin_cit)
prop.table(freq_table)
```
```{r}
cluster_latin_cit <- glm(deported ~ criminality + cluster +latin_cit:cluster+ gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his +  avg_p_nc  + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_latin_cit, file="cluster_latin_cit_bin.rds")
```

```{r}
summary(cluster_latin_cit)
```

```{r}
cluster_latin_cit <- readRDS("models/cluster_latin_cit_bin.rds")
```

#### Get coefficients for Latin citizenship effects at cluster level
```{r}

coefs <- summary(cluster_latin_cit_bin)$coefficients

# Extract main effects for each Latin citizenship category
lat_fixed <- coefs["latin_cit", "Estimate"]

# Extract interaction terms for each category
cluster_int_lat<- coefs[grep("latin_cit$", rownames(coefs)), , drop = FALSE]


# Set non-significant interactions to 0 for each category
cluster_int_lat[cluster_int_lat[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0


# Create interaction dataframes for each category
lat_df <- data.frame(
  cluster = sub("cluster", "", sub(":latin_cit$", "", rownames(cluster_int_lat))),
  cluster_int_lat = cluster_int_lat[, "Estimate"]
)


# Get all unique clusters
all_clusters <- unique(cluster_office_df$cluster)

# Combine all Latin citizenship effects
lat_interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(lat_df, by = "cluster") |>
  mutate(
    # Fixed effects (same for all clusters)
    lat_fixed = lat_fixed,
    # Interaction effects (vary by cluster, 0 if missing/non-sig)
    cluster_int_lat = replace_na(cluster_int_lat, 0),
    clstr_lat_int_sig = ifelse(clstr_asi_int == 0, 1, 0)

  )

# Calculate cluster-specific effects for each category
lat_interaction_df <- lat_interaction_df |>
  mutate(
    cluster_lat_effect = lat_fixed + cluster_int,

  )

# Join to main dataframe
cluster_office_df <- cluster_office_df |>
  left_join(lat_interaction_df, by = "cluster")
```







### Transform interaction terms into percent change
```{r}
head(cluster_office_df)
```



```{r}
cluster_office_df_transf <- cluster_office_df |>
  mutate(cluster_hisp_marg_10pt = cluster_hisp_marg * 0.10,
         cluster_asi_marg_10pt = cluster_asi_marg * 0.10) |> # for 10% increase interpretation
  mutate(across(
    .cols = c(crim_marg, aor_crim_marg, cluster_hisp_marg_10pt, cluster_asi_marg_10pt, cluster_mexico_marg, cluster_samerica_marg, cluster_central_marg),
    .fns = ~ (exp(.) - 1) * 100,
    .names = "{.col}_pctdiff"
  )) 

```

## Export coefficients

```{r}
saveRDS(cluster_office_df_transf, file = "mapping_coefs.rds") 
```

```{r}
st_write(cluster_office_df_transf, "data/mapping_coefs.gpkg")
```


### AIC model comparision
The model interacting cluster with criminality status has the lowest AIC
```{r}
print(AIC(cluster_reg_crim))
print(AIC(cluster_age))
print(AIC(cluster_hisp))
print(AIC(cluster_asian))
print(AIC(cluster_lat_am))
```

# GWR with aggregated offices
My computer crashes every time I try to compute the distance matrix because there are so many observations. So, my plan is to aggregate the data by office, but assign weights to offices with more datapoints. 

```{r}
# aggregate by offices
office_agg <- cnty_clus_offices |>
  group_by(office) |>
  reframe(prop_latin_cit = mean(latin_cit == 1),
         prop_crim_or_pending = mean(criminality == "1 Convicted Criminal" | criminality == "2 Pending Criminal Charges"),
         prop_female = mean(gender == "Female"),
         prop_male = mean(gender == "Male"),
         avg_prop_asn = mean(avg_p_asn),
         avg_prop_his = mean(avg_p_his),
         avg_prop_blk = mean(avg_p_blk),
         avg_prop_nc = mean(avg_p_nc),
         avg_pop_sqm = mean(avg_pop_sqm),
         avg_age = mean(age), 
         prop_deported = mean(deported == 1),
         num_encounters = num_encounters,
         office = office,
         aor = aor,
         geom = st_union(geom)) |>
   distinct(office, .keep_all = TRUE)

office_agg_sf <- st_as_sf(
  office_agg,
  sf_column_name = "geom",
  crs = st_crs(cnty_clus_offices)  
)
```



Note: summary statistics won't really make sense here because of the unequal distribution of encounters per office (weighting would need to be done first)

```{r}
office_agg |>
  summarize(min(num_encounters),
            max(num_encounters),
            mean(num_encounters),
            median(num_encounters),
            sd(num_encounters))
```


```{r}
office_agg_sp <- as(office_agg_sf, "Spatial")

bw <- bw.gwr(prop_deported ~ prop_latin_cit + prop_crim_or_pending + prop_female + avg_prop_asn + avg_prop_his + avg_prop_nc +avg_pop_sqm,
             data = office_agg_sp,
             approach = "CV",
             kernel = "bisquare",
             adaptive=TRUE)
```
## Fit OLS
```{r}
# ols first
ols_fit <- lm(prop_deported ~ prop_latin_cit + prop_crim_or_pending + prop_female + avg_prop_asn + avg_prop_his  + avg_prop_nc + num_encounters,
             data = office_agg_sp)

summary(ols_fit)
```

### Moran's I
```{r}
office_agg_sp$lm_resids <- ols_fit$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. 
```{r}
office_agg_sf <- st_as_sf(office_agg_sp)
office_coords <- st_coordinates(office_agg_sf)


w1 <- 1/as.matrix(st_distance(office_agg_sf,
                              office_agg_sf))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(office_agg_sp$lm_resids, mat2listw(w1, style="W"))
```


```{r}
# replaced population density with number of encounters
gwr_fit <- gwr.basic(prop_deported ~ prop_crim_or_pending + prop_latin_cit + prop_female + avg_prop_asn + avg_prop_his  + avg_prop_nc + num_encounters,
             data = office_agg_sp,
             bw = bw,
             kernel = "bisquare",
             adaptive=TRUE)

gwr_results <- st_as_sf(gwr_fit$SDF)
gwr_fit 
```



```{r}
# alpha* adjustment

#effective number of parameters
pe <- 20.84718 
n <- nrow(office_agg_sp)
p <- 9

alpha_star <- 0.05/(1+pe - pe/(n*p))
alpha_star
df <- n - p - 1
t_star <- qt(1-alpha_star, df = df)
t_star 
```


```{r}
# assign significance
gwr_results <- gwr_results |>
  mutate(sig_crim = ifelse(abs(prop_crim_or_pending_TV) > t_star, prop_crim_or_pending, 0),
         sig_latin_cit = ifelse(abs(prop_latin_cit_TV) > t_star, prop_latin_cit, 0),
         sig_his = ifelse(abs(avg_prop_his_TV) > t_star,avg_prop_his, 0),
         sig_asn = ifelse(abs(avg_prop_asn_TV) > t_star, avg_prop_asn, 0),
         sig_female = ifelse(abs(prop_female_TV)> t_star, prop_female, 0))

hist(gwr_results$sig_crim)
hist(gwr_results$sig_latin_cit)
hist(gwr_results$sig_his)
hist(gwr_results$sig_asn)
hist(gwr_results$sig_female)

```

```{r}
# basic plot
gwr_results <- st_transform(gwr_results, st_crs(cnty_clus_offices))

# plot local coefficients as quintiles
# criminality
q_criminal <- quantile(gwr_results$sig_crim,
                       probs = seq(0, 1, 0.1),
                       na.rm=TRUE)

gwr_results$coef_crim_q <- case_when(
  gwr_results$sig_crim == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_crim,
                          breaks = unique(q_criminal),
                          include.lowest = TRUE))
)

plot_crim <- ggplot(gwr_results) +
  geom_sf(aes(color = coef_crim_q), linewidth = 0.2) +
  scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_crim, 
                                 breaks = unique(q_criminal), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size = 12) + 
  ggtitle("Criminal/Pending Criminal Charges Coef. (quintiles)")

plot_crim

```


```{r}
# 100% significant: space maters
# being from central america
# criminality
q_latin_cit <- quantile(gwr_results$sig_latin_cit,
                       probs = seq(0, 1, 0.2),
                       na.rm=TRUE)

gwr_results$coef_latin_cit_q <- case_when(
  gwr_results$sig_latin_cit == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_latin_cit,
                          breaks = unique(q_latin_cit),
                          include.lowest = TRUE))
)

plot_latin_cit <- ggplot(gwr_results) +
  geom_sf(aes(color = coef_latin_cit_q), linewidth = 0.2) +
  scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_latin_cit, 
                                 breaks = unique(q_latin_cit), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size = 12) + 
  ggtitle("Citizenship from a Latin American Country Coef. (quintiles)")

plot_latin_cit

```



```{r}
q_his <- quantile(gwr_results$avg_prop_his,
                       probs = seq(0, 1, 0.2),
                       na.rm=TRUE)

gwr_results$coef_his_q <- case_when(
  gwr_results$sig_his == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_his,
                          breaks = unique(q_his),
                          include.lowest = TRUE))
)

plot_his<- ggplot(gwr_results) +
  geom_sf(aes(color = coef_his_q), linewidth = 0.2) +
 scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_his, 
                                 breaks = unique(q_his), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size= 12) + 
  theme_bw() 
  ggtitle("Local Hispanic/Latinx Proportion Coef. (quintiles)")

plot_his
```

```{r}
q_asn <- quantile(gwr_results$avg_prop_asn,
                       probs = seq(0, 1, 0.2),
                       na.rm=TRUE)

gwr_results$coef_asn_q <- case_when(
  gwr_results$sig_asn == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_asn,
                          breaks = unique(q_asn),
                          include.lowest = TRUE))
)

plot_asn<- ggplot(gwr_results) +
  geom_sf(aes(color = coef_asn_q), linewidth = 0.2) +
 scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_asn, 
                                 breaks = unique(q_asn), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size= 12) + 
  theme_bw() 
  ggtitle("Local Asian Proportion Coef. (quintiles)")

plot_asn
```

```{r}
q_female <- quantile(gwr_results$sig_female,
                       probs = seq(0, 1, 0.2),
                       na.rm=TRUE)

gwr_results$coef_female_q <- case_when(
  gwr_results$sig_female == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_female,
                          breaks = unique(q_female),
                          include.lowest = TRUE))
)

plot_prop_female<- ggplot(gwr_results) +
  geom_sf(aes(color = coef_female_q), linewidth = 0.2) +
  scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_female, 
                                 breaks = unique(q_female), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size= 12) + 
  theme_bw() 
  ggtitle("Proportion Female Coef. (quintiles)")

plot_prop_female
```





# Stuff not included in paper
## Get predicted probabilities of deportation from cluster:criminality regression model
```{r}
pred_odds <- exp(predict(cluster_reg_crim))
pred_probs <- pred_odds / (pred_odds + 1)

cnty_clus_offices$pred_prob <- pred_probs
```

```{r}
cnty_clus_offices |>
  select(deported, pred_prob) |> 
  st_drop_geometry() |>
  slice_sample(n=10)
```


### Do train/test split to get confusion matrix, F1 score, and ROC curve
```{r}
set.seed(888)


cnty_clus_offices_older <- cnty_clus_offices |>
  filter(evnt_dt < as.Date("2025-04-27"))

# create train-test split, createDataPartition already shuffles your data
split_idx <- createDataPartition(cnty_clus_offices_older$deported, p = 0.8, list=FALSE)

encounters_train = cnty_clus_offices[split_idx, ]

encounters_test = cnty_clus_offices[-split_idx, ]



```

```{r}
cluster_reg_train <- glm(deported ~ criminality +criminality:cluster+ gender + age + region + fnl_prg + avg_p_blk + avg_p_asn  + avg_p_his + avg_pop_sqm, 
               family = binomial,
               data = encounters_train)
```


```{r}
saveRDS(cluster_reg_train, file="cluster_reg_train.rds")
```

```{r}
cluster_reg_train<- readRDS("cluster_reg_train.rds")
```


Get predictions from test set
```{r}
pred_probs <- predict(cluster_reg_train, newdata = encounters_test, type = "response")
```

ROC curve to decide where to place the "threshold" for classification
```{r}
roc_curve <- roc(response = encounters_test$deported, predictor = pred_probs)

plot(roc_curve, main = "ROC Curve", print.auc=TRUE,
     xlim = c(0, 1))
auc(roc_curve)
```


```{r}
# convert probabilities to predicted class
# TODO: ask about this
pred_dep <- ifelse(pred_probs >= 0.50, "1", "0")
pred_dep <- factor(pred_dep, levels= levels(encounters_test$deported))

# create confusion matrix
confusionMatrix(
  data = pred_dep,
  reference = encounters_test$deported,
  positive = "1" # well, not actually "positive"
)

```
```{r}
# proportion of deport predictions that were actually correct 
precision <- posPredValue(pred_dep, encounters_test$deported, positive = "1")

# proportion of correctly predicted deported cases
recall <- sensitivity(pred_dep, encounters_test$deported, positive = "1")

f1 <- (2*precision * recall) / (precision + recall)

print(precision)
print(recall)
print(f1)
```


```{r}
hist(pred_probs)
```

