---
title: "Modeling Framework"
format: html
editor: source
---
```{r}
library(tidyverse)
library(sf)
library(VGAM)
library(lubridate)
library(lme4)
library(spdep)
library(dbscan)
library(car)
library(corrplot)
library(GWmodel)
library(raster)
library(spdep)
library(paletteer)
```

Read in clustered data
```{r}
cnty_clus_offices <- st_read("data/cnty_clus_offices.shp")

cnty_clus_offices <- cnty_clus_offices |>
  rename("avg_p_a" = "avg_p_s",
         "avg_pop_sqm" = "avg_pp_") |>
  drop_na(crmnlty, gender, age, region, avg_p_b, avg_p_a, avg_p_h, avg_pop_sqm)
```



### Check for multicollinearity
```{r}
numeric_vars <- cnty_clus_offices |>
  group_by(office) |>
  select(c(avg_p_b, avg_p_a, avg_p_h, avg_p_n, avg_pop_sqm)) |>
  drop_na |>
  st_drop_geometry()
```

```{r}
corr_mat <- cor(numeric_vars)
corrplot(corr_mat, type="upper", method="ellipse")
```
```{r}
corr_mat
```

## Basic logistic regression
```{r}
log_reg <-glm(deportd ~ crmnlty + gender + age + region + avg_p_b + avg_p_a + avg_p_h + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(log_reg , file="log_reg1.rds")
```

```{r}
summary(log_reg )
```

```{r}
cnty_clus_offices$log_resids <- log_reg$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. We conclude that there is clustering behavior of the residuals. 
```{r}
offices_grouped <- cnty_clus_offices |>
  group_by(office) |>
  mutate(avg_log_resid = mean(log_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_log_resid, mat2listw(w1, style="W"))
```


## Interacting criminality with cluster 
Cluster reference group: 167
```{r}

cluster_reg <- glm(deportd ~ crmnlty + crmnlty:clstr_f + gender + age + region + avg_p_b + avg_p_a + avg_p_h + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_reg, file="cluster_reg1.rds")
```

```{r}
summary(cluster_reg)
```

Get coefficients for the interaction term for noncriminals
```{r}
cluster_interaction_coefs_3 <- coef(cluster_reg)[grep("crmnlty3 Other Immigration Violator:", names(coef(cluster_reg)))]
```

Note: The interaction term tests whether the effect of being a noncriminal on the outcome differs in cluster i compared to the reference cluster (167).
```{r}

interaction_df <- data.frame(
  term = names(cluster_interaction_coefs_3),
  clstr_crim_3_int_coef = as.numeric(cluster_interaction_coefs_3)
) |>
  mutate(clstr_f = sub(".*clstr_f", "", term)) |>
  select(-term)

#clusters_grouped <- cnty_clus_offices |>
  #group_by(clstr_f, office, geometry)

clusters_interactions <- cnty_clus_offices |>
  left_join(interaction_df, by = "clstr_f")
```

## Interacting criminality with aor

### exploration: calculate the proportion noncriminal per aor
```{r}
offices_clean |>
  group_by(aor)|>
  summarize(prop_noncrim = mean(criminality == "3 Other Immigration Violator"))
```


```{r}

aor_reg <- glm(deportd ~ crmnlty + crmnlty:aor + gender + age + region + avg_p_b + avg_p_a + avg_p_h + avg_pop_sqm, 
               family = binomial,
               data = cnty_clus_offices)


saveRDS(aor_reg, file="aor_reg1.rds")
```

```{r}
summary(aor_reg)
```

Get coefficients for the interaction term
```{r}
aor_interaction_coefs <- coef(aor_reg)[grep("criminality", names(coef(aor_reg)))]
```

```{r}
aor_interaction_coefs
```


### Use distance to Mexican border as the interaction term
```{r}
office_dists <- read_csv("data/dists_to_border.csv")
```
```{r}
office_border_dists <- office_dists |>
  rename("avg_p_a" = "avg_p_s",
         "avg_pop_sqm" = "avg_pp_")
```


```{r}
dists_reg <- glm(deportd ~ crmnlty:HubDist + gender + age + region + avg_p_b + avg_p_a + avg_p_h + avg_p_n + avg_pop_sqm, 
               family = binomial,
               data = office_border_dists)

saveRDS(dists_reg, file="dists_reg.rds")
```


```{r}
summary(dists_reg)
```

### Hotspot analysis for number of encounters per office
Getis-Ord Gi* Hotspot with 10km Grid
```{r}
num_encounters_agg <- cnty_clus_offices|>
  group_by(office) |>
  summarize(num_enc = n())
```

```{r}
counties <- st_read("data/county_info_pr.shp")
```


```{r}
# 80km grid ~ about 50 miles
pixelsize = 80000
box = round(extent(counties)/pixelsize) * pixelsize
template = raster(box, crs = "ESRI:102010",
                  nrows = (box@ymax - box@ymin) / pixelsize,
                  ncols = (box@xmax - box@xmin) / pixelsize)
```

```{r}
getisraster=rasterize(num_encounters_agg, template, field = 'num_enc')
getisgrid = rasterToPolygons(getisraster)
```

Create list of neighbors
```{r}
neighbors = poly2nb(getisgrid)
weighted_neighbors = nb2listw(neighbors, zero.policy = T)

```
Perform the local G analysis
```{r}
getisgrid$hotspot = as.vector(localG(getisgrid$layer, weighted_neighbors))
```

PLot
```{r}
breaks_hs = (-Inf, -1.96, -1, 1, 1.96, Inf)



palette_hs = paletteer_c("ggthemes::Red-Blue-White Diverging", 5, direction=-1)

col_hs=palette_hs[cut(getisgrid$hotspot, breaks_hs)]
plot(counties['cnty_fp'], col="transparent", border="gray50", reset=FALSE, main = "Hotspots")
plot(getisgrid, col=col_hs, border=NA, add=T)
```

Compare with actual num encounters
```{r}
breaks_num_enc = quantile(num_encounters_agg$num_enc)
num_enc_palette = paletteer_c("ggthemes::Red", 5)
num_enc_col = palette[cut(num_encounters_agg$num_enc, breaks_num_enc )]
```

```{r}
getis_sf <- st_as_sf(getisgrid)
counties_sf <- st_as_sf(counties)
num_enc_sf <- st_as_sf(num_encounters_agg)

getis_sf <- st_join(getis_sf, num_enc_sf, left = TRUE)

breaks_hs_q = quantile(getis_sf$hotspot, na.rm=TRUE)
getis_sf$hotspot_cat <- cut(getis_sf$hotspot, breaks = breaks_hs_q)



ggplot() + 
  geom_sf(data=counties_sf, fill=NA, color="gray50", linewidth=0.3) +
  geom_sf(data=getis_sf, aes(fill=hotspot_cat, color = NA) +
            scale_fill_manual(values=palette_hs, name = "Hotspot (Gi&)") +
            coord_sf())
```


### Stuff that prob won't work
GWR (aggregated to the office level)
```{r}
office_agg <- cnty_clus_offices |>
  group_by(office) |>
  reframe(prop_noncrim = mean(crmnlty == "3 Other Immigration Violator"),
          prop_deported = mean(deportd == 1),
            avg_p_b = first(avg_p_b),
            avg_p_a = first(avg_p_a),
            avg_p_h = first(avg_p_h),
            avg_p_n = first(avg_p_n),
            avg_pop_sqm = first(avg_pop_sqm),
          geometry=first(geometry)) |>
  st_as_sf()
```

Create distance matrix
```{r}
office_agg_sf <- as(office_agg, "Spatial")
office_dists <- gw.dist(dp.locat=coordinates(office_agg_sf))
```

Bandwidth selection
bw_gwr <- bw.gwr(
  log(homicides) ~ indice,
  data=muni_sp,
  adaptive=TRUE,
  dMat = muni_dists,
  approach="CV"
)
```{r}
bw_gwr <- bw.gwr(
  prop_deported ~ prop_noncrim + avg_p_b + avg_p_a + avg_p_h + avg_p_n + avg_pop_sqm,
  data=office_agg_sf,
  adaptive=TRUE,
  dMat=office_dists,
  approach="CV"
)
```


Spatial Hierarchal Modeling

Hierarchical model with random intercepts and a random slope for criminality for each office and or
```{r}
# sample from the dataset bc it takes too long to run
#set.seed(42069)
#rand_sample <- offices_clustered[sample(nrow(offices_clustered), size = 5000), ]

glmer_reg <- glmer(
  deported ~ criminality + gender + scale(age) + region +
    (1 | cluster_final) + (0 + criminality | cluster_final),
  family = binomial,
  data = offices_clustered
)

```

```{r}
rand_sample_df <- rand_sample |>
  st_drop_geometry() |>
  group_by(office)|>
  count()

rand_sample_df
```

```{r}
offices_clean |>
  st_drop_geometry() |>
  group_by(office)|>
  count() |>
  arrange(desc(n))

```



```{r}
summary(glmer_reg)
```

Random intercepts and slopes
```{r}
rand_effects_office <- ranef(glmer_reg)$office
rand_effects_aor <- ranef(glmer_reg$`aor:office`)

```

Map the office-level slopes
```{r}
rand_sample <- rand_sample |>
  group_by(office) |>
  left_join(rand_effects, by = c("office" = rownames(rand_sample)))

# convert to sf
rand_sample_sf <- st_as_sf(rand_sample, crs="ESRI:102010")
```

