---
title: "Modeling Framework"
format: html
editor: source
---
```{r}
library(tidyverse)
library(sf)
library(VGAM)
library(lubridate)
library(lme4)
library(spdep)
library(car)
library(corrplot)
library(paletteer)
library(MASS)
library(glue)
library(spatialreg)
select <- dplyr::select
```

Read in clustered data
```{r}
cnty_clus_offices <- readRDS("data/cnty_clus_offices.rds")


cnty_clus_offices <- cnty_clus_offices |>
  drop_na(criminality, gender, age, region)


cnty_clus_offices$deported <- factor(cnty_clus_offices$deported, levels = c(0,1))

cnty_clus_offices$region <- factor(cnty_clus_offices$region)

cnty_clus_offices$cluster <- factor(cnty_clus_offices$cluster)

cnty_clus_offices$gender <- factor(cnty_clus_offices$gender)

cnty_clus_offices$criminality <- factor(cnty_clus_offices$criminality)

cnty_clus_offices$criminality <- relevel(cnty_clus_offices$criminality, ref = "3 Other Immigration Violator")

cnty_clus_offices$cluster <- relevel(cnty_clus_offices$cluster, ref = "43")


# split into Latin American / Non-Latin American citizenship
latin_am_regions = c("Central America", "Caribbean", "Mexico", "South America")

cnty_clus_offices <- cnty_clus_offices |>
  mutate(latin_cit = case_when(region == "Mexico" ~ 1,
                                  region == "South America" ~ 1,
                                  region == "Central America" | region == "Caribbean" ~ 1,
                                  !(region %in% latin_am_regions) ~ 0)) 



trump_enc <- cnty_clus_offices |>
  filter(trump == 1)

biden_enc <- cnty_clus_offices |>
  filter(trump == 0)

```

```{r}
count(cnty_clus_offices |> filter(cluster %in% c("25","106")), trump)

```
```{r}
trump_enc |>
  filter(cluster %in% c("25","106"))
```


### Check for multicollinearity
```{r}
numeric_vars <- cnty_clus_offices |>
  group_by(office) |>
  select(c(avg_p_blk, avg_p_asn , avg_p_his, avg_p_nc, avg_pop_sqm)) |>
  drop_na() |>
  st_drop_geometry()
```

None of correlations are above 0.8, so I can keep all variables in the model
```{r}
corr_mat <- cor(numeric_vars)
corrplot(corr_mat, type="upper", method="ellipse")
```
```{r}
corr_mat
```

## Basic logistic regression

### Trump
```{r}
trump_enc_sub <- trump_enc |>
  drop_na(c(criminality, gender, age, latin_cit, avg_p_blk, avg_p_asn, avg_p_his, avg_p_nc,avg_pop_sqm, tps))

log_reg_trump <-glm(deported ~ criminality + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_p_nc + avg_pop_sqm + tps, 
               family = binomial,
               data = trump_enc_sub)

saveRDS(log_reg_trump , file="models/log_reg_trump.rds")
```

```{r}
#log_reg_trump <- readRDS("models/log_reg_trump.rds")
```

```{r}
summary(log_reg_trump)
```

### Biden
```{r}
biden_enc_sub <- biden_enc |>
  drop_na(c(criminality, gender, age, latin_cit, avg_p_blk, avg_p_asn, avg_p_his, avg_p_nc,avg_pop_sqm, tps))
 
log_reg_biden <-glm(deported ~ criminality + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_p_nc + avg_pop_sqm + tps, 
               family = binomial,
               data = biden_enc)

saveRDS(log_reg_biden , file="models/log_reg_biden.rds")

#log_reg_biden <- readRDS("models/log_reg_biden.rds")
#summary(log_reg_biden)
```

## Administration 
```{r}
log_reg_admin <-glm(deported ~ criminality + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_pop_sqm + tps + trump, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(log_reg_admin, "models/log_reg_admin")
summary(log_reg_admin)
```



### Moran's I on basic logistic regression
Biden
```{r}
biden_enc_sub$log_resids <- log_reg_biden$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated.
```{r}
offices_grouped <- biden_enc_sub |>
  group_by(office) |>
  mutate(avg_log_resid = mean(log_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_log_resid, mat2listw(w1, style="W"))
```


Trump
```{r}
trump_enc_sub$log_resids <- log_reg_trump$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. However, our p-value is pretty close to 0.05 (0.06). 
```{r}
offices_grouped <- trump_enc_sub |>
  group_by(office) |>
  mutate(avg_log_resid = mean(log_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_log_resid, mat2listw(w1, style="W"))
```

Biden
```{r}
biden_enc$log_resids <- log_reg_biden$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. We conclude that there is clustering behavior of the residuals. 
```{r}
offices_grouped <- biden_enc |>
  group_by(office) |>
  mutate(avg_log_resid = mean(log_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_log_resid, mat2listw(w1, style="W"))
```



## Interacting criminality with cluster 
Overall
```{r}
crim_reg <- glm(deported ~ criminality + cluster + criminality:cluster + gender + age + latin_cit + tps, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(crim_reg, file="models/crim_reg.rds")
```

```{r}
get_overall_clus_effect_summary <- function(model, key_var) {
  coefs <- summary(model)$coefficients
  rownames <- rownames(coefs)
  
  # main effect
  main_effect <- coefs[key_var, "Estimate"]
  
  # pattern for both orders
  int_pattern <- paste0("^cluster([0-9]+):", key_var, "$|^", key_var, ":cluster([0-9]+)$")
  
  # extract matching rows
  int_rows <- grep(int_pattern, rownames, value = TRUE)
  int_coefs <- coefs[int_rows, , drop = FALSE]
  
  # set non-significant to 0
  int_coefs[int_coefs[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0
  
  cluster_effect <- main_effect + int_coefs
  
   quants <- paste(round(quantile(cluster_effect, probs = seq(0, 1, .2)), 4), collapse = ", ")
  jenks  <- paste(round(getJenksBreaks(cluster_effect, 5), 4), collapse = ", ")
  sd_val <- round(sd(cluster_effect), 4)
  
  return(glue("quantiles: {quants}\njenks: {jenks}\nsd: {sd_val}"))

}



get_overall_clus_effect_summary(crim_reg, "criminality1 Convicted Criminal")

```


Cluster reference group: 167
```{r}
crim_trump <- glm(deported ~ criminality + cluster + criminality:cluster + gender + age + latin_cit + tps, 
               family = binomial,
               data = trump_enc)

saveRDS(crim_trump, file="models/crim_trump.rds")
```

```{r}
crim_trump<- readRDS("models/crim_trump.rds")
summary(crim_trump)
```


```{r}
crim_biden <- glm(deported ~ criminality + cluster + criminality:cluster + gender + age + latin_cit + tps, 
               family = binomial,
               data = biden_enc)

saveRDS(crim_biden, file="models/crim_biden.rds")
```

```{r}
crim_biden<- readRDS("models/crim_biden.rds")
summary(crim_biden)
```




### Moran's I on cluster:criminality model
Biden
```{r}
biden_enc$crim_biden_resids <- crim_biden$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- biden_enc|>
  group_by(office) |>
  mutate(avg_crim_biden_resids = mean(crim_biden_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_crim_biden_resids, mat2listw(w1, style="W"))
```
Trump
```{r}
trump_enc$crim_trump_resids <- crim_trump$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- trump_enc|>
  group_by(office) |>
  mutate(avg_crim_trump_resids = mean(crim_trump_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_crim_trump_resids, mat2listw(w1, style="W"))
```

## Interaction LatinAmerica:cluster

```{r}
latin_trump <- glm(deported ~ criminality + cluster +latin_cit:cluster+ gender + age + latin_cit + tps, 
               family = binomial,
               data = trump_enc)

saveRDS(latin_trump, file="models/latin_trump.rds")
```

```{r}
latin_trump <- readRDS("models/latin_trump.rds")
summary(latin_trump)
```

`


```{r}
latin_biden <- glm(deported ~ criminality + cluster +latin_cit:cluster+ gender + age + latin_cit + tps, 
               family = binomial,
               data = biden_enc)

saveRDS(latin_biden, file="models/latin_biden.rds")


```
```{r}
summary(latin_biden)
```

```{r}


# define neighbors (example: based on coordinates)
coords <- st_coordinates(biden_enc)
nb <- dnearneigh(coords,
                 d1 = 0,
                 d2 = 160000, # 160 km
                 longlat = FALSE)

# create spatial weights
lw <- nb2listw(nb, style = "W")

latin_biden_spatial <- lagsarlm(deported ~ criminality + cluster +
                               latin_cit:cluster + gender + age +
                               latin_cit + tps,
                               data = biden_enc,
                               listw = lw)
```


```{r}
summary(latin_biden$coefficients)

```

```{r}
robust_latin_biden
```


```{r}
latin_biden <- readRDS("models/latin_biden.rds")
summary(latin_biden)
```

### Moran's I on cluster:latin model
Biden
```{r}
biden_enc$latin_biden_resids <- latin_biden$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. So cannot be fully trusted. 
```{r}
offices_grouped <- biden_enc|>
  group_by(office) |>
  mutate(avg_latin_biden_resids = mean(latin_biden_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_latin_biden_resids, mat2listw(w1, style="W"))
```
Trump
```{r}
trump_enc$latin_trump_resids <- latin_trump$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- trump_enc|>
  group_by(office) |>
  mutate(avg_latin_trump_resids = mean(latin_trump_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_latin_trump_resids, mat2listw(w1, style="W"))
```

## Interaction tps:cluster

```{r}
tps_trump <- glm(deported ~ criminality + cluster +tps:cluster+ gender + age + latin_cit + tps, 
               family = binomial,
               data = trump_enc)

saveRDS(tps_trump, file="models/tps_trump.rds")
```

```{r}
tps_trump <- readRDS("models/tps_trump.rds")
summary(tps_trump)
```

```{r}
tps_biden <- glm(deported ~ criminality + cluster +tps:cluster+ gender + age + latin_cit + tps, 
               family = binomial,
               data = biden_enc)

saveRDS(tps_biden, file="models/tps_biden.rds")
```

```{r}
tps_biden<- readRDS("models/tps_biden.rds")
summary(tps_biden)
```

### Moran's I on cluster:tps model
Biden
```{r}
biden_enc$tps_biden_resids <- tps_biden$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. 
```{r}
offices_grouped <- biden_enc|>
  group_by(office) |>
  mutate(avg_tps_biden_resids = mean(tps_biden_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_tps_biden_resids, mat2listw(w1, style="W"))
```
Trump
```{r}
trump_enc$latin_trump_resids <- latin_trump$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- trump_enc|>
  group_by(office) |>
  mutate(avg_latin_trump_resids = mean(latin_trump_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_latin_trump_resids, mat2listw(w1, style="W"))
```


## Extract coefficients from all models
```{r}
model_names <- c('crim_trump', 'crim_biden', 'latin_trump', 'latin_biden', 'tps_trump', 'tps_biden')
models <- mget(model_names)
key_vars <- c("criminality1 Convicted Criminal", "criminality1 Convicted Criminal", "latin_cit", "latin_cit", "tps", "tps")

coef_extraction <- function(model, key_var, data, model_name, robust=NULL) {
  # use robust SE if provided, otherwise use standard summary
  if (!is.null(robust)) {
    coefs <- robust
  } else {
    coefs <-summary(model)$coefficients
  }
  
  rownames <- rownames(coefs)
  
  # main effect
  main_effect <- coefs[key_var, "Estimate"]
  
  # pattern for both orders
  int_pattern <- paste0("^cluster([0-9]+):", key_var, "$|^", key_var, ":cluster([0-9]+)$")
  
  # extract matching rows
  int_rows <- grep(int_pattern, rownames, value = TRUE)
  int_coefs <- coefs[int_rows, , drop = FALSE]
  
  # set non-significant to 0
  int_coefs[int_coefs[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0
  
  # extract cluster number regardless of order
  cluster_num <- ifelse(
    grepl(paste0("^cluster([0-9]+):", key_var, "$"), rownames(int_coefs)),
    sub(paste0("^cluster([0-9]+):", key_var, "$"), "\\1", rownames(int_coefs)),
    sub(paste0("^", key_var, ":cluster([0-9]+)$"), "\\1", rownames(int_coefs))
  )
  
  if (is.factor(data$cluster)) {
    cluster_num <- factor(cluster_num, levels = levels(data$cluster))
  } else {
    cluster_num <- as(cluster_num, class(data$cluster))
  }
  
  int_df <- data.frame(
    cluster = factor(cluster_num, levels = levels(factor(data$cluster))),
    clstr_int = int_coefs[, "Estimate"]
  )
  
  # get all unique clusters
  clusters <- unique(data$cluster)
  
  # combine with main effect
  interaction_df <- data.frame(cluster = clusters) %>%
    left_join(int_df, by = "cluster") %>%
    mutate(
      main_effect = main_effect,
      clstr_int = ifelse(is.na(clstr_int), 0, clstr_int),
      cluster_effect = main_effect + clstr_int
    )
  
  # rename columns
  interaction_df <- interaction_df %>%
    rename(
      !!paste0(model_name, "_main_effect") := main_effect,
      !!paste0(model_name, "_clstr_int") := clstr_int,
      !!paste0(model_name, "_cluster_eff") := cluster_effect
    )
  
  return(interaction_df)
}
```

```{r}
coef_df_list <- mapply(
  coef_extraction,
  model = models,
  key_var = key_vars,
  model_name = model_names,
  MoreArgs = list(data=cnty_clus_offices),
  SIMPLIFY = FALSE
)



coefs_df <- Reduce(
  left_join,
  coef_df_list
)
```

```{r}
max(coefs_df$latin_biden_cluster_eff)
```



```{r}
coefs_office_joined <- cnty_clus_offices |>
  group_by(office) |>
  slice(1) |>
  select(office, cluster, geom) |> 
  left_join(coefs_df, by="cluster")

head(coefs_office_joined)
```


### Transform interaction terms into percent change
```{r}
coefs_df_transf <- coefs_office_joined |>
  mutate(across(
    .cols = c(crim_trump_cluster_eff, crim_biden_cluster_eff, latin_trump_cluster_eff, latin_biden_cluster_eff, tps_trump_cluster_eff, tps_biden_cluster_eff),
    .fns = ~ (exp(.) - 1) * 100,
    .names = "{str_replace(.col, '_cluster_eff', '')}_pctdiff"
  )) |>
  select(c(office, cluster, crim_trump_pctdiff, crim_biden_pctdiff, latin_trump_pctdiff, latin_biden_pctdiff, tps_trump_pctdiff, tps_biden_pctdiff))

```

```{r}
max(coefs_df_transf$latin_biden_pctdiff)
```


```{r}

coefs_df_transf |>
 filter(if_any(everything(), is.na)) |>
  distinct(cluster, .keep_all = TRUE)
```

```{r}
nrow(coefs_df_transf)
```



## Export coefficients

```{r}
saveRDS(coefs_df_transf, file = "mapping_coefs_admin.rds") 
saveRDS(cnty_clus_offices, file = "cnty_clus_offices.rds") 
```


```{r}
st_write(coefs_df_transf, "data/mapping_coefs_admin.gpkg")
```


### AIC model comparision
```{r}
print(AIC(log_reg_trump))
print(AIC(crim_trump))
print(AIC(latin_trump))
print(AIC(tps_trump))
```

```{r}
print(AIC(log_reg_biden))
print(AIC(crim_biden))
print(AIC(latin_biden))
print(AIC(tps_biden))
```

## Compare noncriminal Biden vs. Trump 

```{r}
set.seed(522)
# create profile
clusters <- levels(cnty_clus_offices$cluster)
newdata <- data.frame(
  criminality = "3 Other Immigration Violator", # noncriminal
  cluster = clusters, # vary by cluster, then take the mean 
  gender = "Male", # mode
  age = mean(cnty_clus_offices$age),
  latin_cit = 1, # mode
  tps = 0 # mode
  
)

# set factors
newdata$criminality <- factor(
  newdata$criminality,
  levels = levels(cnty_clus_offices$criminality)
)
newdata$cluster <- factor(
  newdata$cluster,
  levels = levels(cnty_clus_offices$cluster)
)
newdata$gender <- factor(
  newdata$gender,
  levels = levels(cnty_clus_offices$gender)
)
newdata$latin_cit <- factor(
  newdata$latin_cit,
  levels = levels(factor(cnty_clus_offices$latin_cit))
)
newdata$tps <- factor(
  newdata$tps,
  levels = levels(factor(cnty_clus_offices$tps))
)


# design matrix
X <- model.matrix(~ criminality + cluster + criminality:cluster + gender + age + latin_cit + tps, data=newdata)
```

```{r}
# simulate for both trump and biden
n_sims <- 10000
sims_trump <- mvrnorm(n_sims, coef(crim_trump), vcov(crim_trump))
sims_biden <- mvrnorm(n_sims, coef(crim_biden), vcov(crim_biden))

# predict for all clusters (each column is a cluster)
linpred_trump <- sims_trump %*% t(X)
probs_trump <- 1/(1+exp(-linpred_trump))

linpred_biden <- sims_biden %*% t(X)
probs_biden <- 1/(1+exp(-linpred_biden))

# take the median across all clusters
trump_median <- apply(probs_trump, 1, median, na.rm =TRUE)
biden_median <- apply(probs_biden, 1, median, na.rm=TRUE)
```

```{r}
mean(trump_median)
mean(biden_median)

diff <- trump_median - biden_median
mean(diff)

# uncertainty
quantile(diff, c(0.025, 0.975))
```

## Compare Latin American Citizenship Biden vs. Trump

```{r}
clusters <- levels(cnty_clus_offices$cluster)
lat_profile <- data.frame(
  criminality = "3 Other Immigration Violator", # noncriminal
  cluster = clusters, # vary by cluster, then take the mean 
  gender = "Male", # mode
  age = mean(cnty_clus_offices$age),
  latin_cit = 1, 
  tps = 0 # mode
  
)

nonlat_profile <- data.frame(
  criminality = "3 Other Immigration Violator", # noncriminal
  cluster = clusters, # vary by cluster, then take the mean 
  gender = "Male", # mode
  age = mean(cnty_clus_offices$age),
  latin_cit = 0, # mode
  tps = 0 # mode
  
)


# set factors
lat_profile$criminality <- factor(
  lat_profile$criminality,
  levels = levels(cnty_clus_offices$criminality)
)
lat_profile$cluster <- factor(
  lat_profile$cluster,
  levels = levels(cnty_clus_offices$cluster)
)
lat_profile$gender <- factor(
  lat_profile$gender,
  levels = levels(cnty_clus_offices$gender)
)
lat_profile$latin_cit <- factor(
  lat_profile$latin_cit,
  levels = levels(factor(cnty_clus_offices$latin_cit))
)
lat_profile$tps <- factor(
  lat_profile$tps,
  levels = levels(factor(cnty_clus_offices$tps))
)

nonlat_profile$criminality <- factor(
  nonlat_profile$criminality,
  levels = levels(cnty_clus_offices$criminality)
)
nonlat_profile$cluster <- factor(
  nonlat_profile$cluster,
  levels = levels(cnty_clus_offices$cluster)
)
nonlat_profile$gender <- factor(
  nonlat_profile$gender,
  levels = levels(cnty_clus_offices$gender)
)
nonlat_profile$latin_cit <- factor(
  nonlat_profile$latin_cit,
  levels = levels(factor(cnty_clus_offices$latin_cit))
)
nonlat_profile$tps <- factor(
  nonlat_profile$tps,
  levels = levels(factor(cnty_clus_offices$tps))
)
```


```{r}
# design matrix
X_lat_profile <- model.matrix(~ criminality + cluster + latin_cit:cluster + gender + age + latin_cit + tps, data=lat_profile)

X_nonlat_profile <- model.matrix(~ criminality + cluster + latin_cit:cluster + gender + age + latin_cit + tps, data=nonlat_profile)

# simulate for both trump and biden
n_sims <- 10000
sims_trump <- mvrnorm(n_sims, coef(latin_trump), vcov(latin_trump))
sims_biden <- mvrnorm(n_sims, coef(latin_biden), vcov(latin_biden))

# predict for all clusters (each column is a cluster)
linpred_trump_lat <- sims_trump %*% t(X_lat_profile)
probs_trump_lat <- 1/(1+exp(-linpred_trump_lat))

linpred_trump_nonlat <- sims_trump %*% t(X_nonlat_profile)
probs_trump_nonlat <- 1/(1+exp(-linpred_trump_nonlat))

linpred_biden_lat <- sims_biden %*% t(X_lat_profile)
probs_biden_lat <- 1/(1+exp(-linpred_biden_lat))

linpred_biden_nonlat <- sims_biden %*% t(X_nonlat_profile)
probs_biden_nonlat <- 1/(1+exp(-linpred_biden_nonlat))

# take the median across all clusters
trump_median_lat <- apply(probs_trump_lat, 1, median, na.rm =TRUE)
trump_median_nonlat <- apply(probs_trump_nonlat, 1, median, na.rm =TRUE)

biden_median_lat <- apply(probs_biden_lat, 1, median, na.rm =TRUE)
biden_median_nonlat <- apply(probs_biden_nonlat, 1, median, na.rm =TRUE)

```



```{r}
mean(trump_median_lat )
mean(biden_median_lat)

diff <- trump_median_lat  - biden_median_lat
mean(diff)

# uncertainty
quantile(diff, c(0.025, 0.975))
```
```{r}
mean(trump_median_nonlat )
mean(biden_median_nonlat)

diff <- trump_median_nonlat  - biden_median_nonlat
mean(diff)

# uncertainty
quantile(diff, c(0.025, 0.975))
```


```{r}
biden_median_lat <- data.frame(biden_median_lat)
biden_median_nonlat <- data.frame(biden_median_nonlat)

trump_median_lat <- data.frame(trump_median_lat)
trump_median_nonlat <- data.frame(trump_median_nonlat)
# histogram of the predictions
biden_median_lat$admin <- "Biden"
biden_median_lat$cit <- "Latin American"
biden_median_nonlat$admin <- "Biden"
biden_median_nonlat$cit <- "Non-Latin American"

trump_median_lat$admin <- "Trump"
trump_median_lat$cit <- "Latin American"
trump_median_nonlat$admin <- "Trump"
trump_median_nonlat$cit <- "Non-Latin American"

biden_median_lat <-biden_median_lat|>
  rename("prob" = "biden_median_lat")

biden_median_nonlat <-biden_median_nonlat|>
  rename("prob" = "biden_median_nonlat")

trump_median_lat <- trump_median_lat|>
  rename("prob" = "trump_median_lat")

trump_median_nonlat <- trump_median_nonlat|>
  rename("prob" = "trump_median_nonlat")

median_probs <- rbind(biden_median_lat, biden_median_nonlat, trump_median_lat, trump_median_nonlat)


```


```{r}
median_probs <- median_probs |>
  mutate(across(cit, ~factor(., levels=c("Non-Latin American", "Latin American"))))
median_probs |>
  ggplot(aes(prob, fill = admin)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 50) + 
  scale_fill_manual(values=c("Biden" = "blue", "Trump" = "red")) + 
  labs(title = "Distribution of Simulated Predicted Probabilities of Deportation by Citizenship",
       fill = "Administration",
       caption = "For a noncriminal 36 year-old male without TPS, with the median taken across clusters (n = 10000).") +
  xlab("Simulated Predicted Probabilities") + 
  ylab("Count") +
  theme(plot.caption = element_text(hjust = 0)) + 
  theme_bw() + 
  facet_wrap(~ cit)

ggsave("visuals_and_tables/latin_comparison.png")

```

