---
title: "Modeling Framework"
format: html
editor: source
---
```{r}
library(tidyverse)
library(sf)
library(VGAM)
library(lubridate)
library(lme4)
library('glmmTMB')
library(spdep)
```

Read in responsible sites data
```{r}
offices0 <- st_read("data/us_ice_office_points.shp")
```

```{r}
this_year <- as.integer(format(Sys.Date(), "%Y"))

offices <- offices0 |>
  rename(office = responsi_1) |> 
  rename(dep_country = departure_) |>
  rename(criminality = encounter_) |>
  rename(cit_country = citizenshi) |>
  rename(aor = responsibl) |> 
  mutate(deported = ifelse(dep_country == "Not applicable", 0, 1),
         age = this_year - birth_year)
```

```{r}
offices <- offices |> 
  mutate(
    cit_country = str_to_title(cit_country),
    region = case_when(
      # --- Canada + Mexico ---
      cit_country == "Canada" ~ "Canada",
      cit_country == "Mexico" ~ "Mexico",
  
      # --- Caribbean (including dependencies) ---
      cit_country %in% c(
        "Jamaica", "Haiti", "Dominican Republic", "Cuba",
        "Puerto Rico", "US Virgin Islands", "British Virgin Islands",
        "Cayman Islands", "Aruba", "Curacao", "Bonaire",
        "Guadeloupe", "Martinique", "Bermuda", "Montserrat",
        "Anguilla", "Saint Martin", "Sint Maarten",
        "Antigua and Barbuda", "Dominica", "Grenada",
        "Saint Lucia", "Saint Vincent and the Grenadines",
        "Trinidad and Tobago", "The Bahamas", "Barbados",
        "Saint Kitts and Nevis", "Turks and Caicos Islands"
      ) ~ "Caribbean",
  
      # --- Central America ---
      cit_country %in% c(
        "Belize", "Guatemala", "Honduras", "El Salvador",
        "Nicaragua", "Costa Rica", "Panama"
      ) ~ "Central America",
  
      # --- South America ---
      cit_country %in% c(
        "Argentina", "Bolivia", "Brazil", "Chile", "Colombia",
        "Ecuador", "Guyana", "Paraguay", "Peru",
        "Suriname", "Uruguay", "Venezuela"
      ) ~ "South America",
  
      # --- Africa ---
      cit_country %in% c(
        "Algeria", "Angola", "Benin", "Botswana", "Burkina Faso",
        "Burundi", "Cabo Verde", "Cameroon", "Central African Republic",
        "Chad", "Comoros", "Congo", "Democratic Republic of the Congo",
        "Djibouti", "Egypt", "Equatorial Guinea", "Eritrea",
        "Eswatini", "Ethiopia", "Gabon", "Gambia", "Ghana",
        "Guinea", "Guinea-Bissau", "Ivory Coast", "Kenya",
        "Lesotho", "Liberia", "Libya", "Madagascar", "Malawi",
        "Mali", "Mauritania", "Mauritius", "Morocco", "Mozambique",
        "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe",
        "Senegal", "Seychelles", "Sierra Leone", "Somalia",
        "South Africa", "South Sudan", "Sudan", "Tanzania",
        "Togo", "Tunisia", "Uganda", "Zambia", "Zimbabwe"
      ) ~ "Africa",
  
      # --- Asia ---
      cit_country %in% c(
        # East Asia
        "China", "Japan", "South Korea", "North Korea", "Mongolia", "Taiwan",
        # Southeast Asia
        "Indonesia", "Thailand", "Vietnam", "Malaysia",
        "Singapore", "Philippines", "Cambodia", "Laos",
        "Myanmar", "Brunei", "Timor-Leste",
        # South Asia
        "India", "Pakistan", "Bangladesh", "Nepal",
        "Sri Lanka", "Bhutan", "Maldives",
        # Central Asia (your rule)
        "Kazakhstan", "Kyrgyzstan", "Tajikistan",
        "Turkmenistan", "Uzbekistan",
        # Middle East (your rule)
        "Saudi Arabia", "United Arab Emirates", "Qatar", "Kuwait",
        "Bahrain", "Oman", "Yemen", "Jordan",
        "Lebanon", "Syria", "Iraq", "Iran",
        "Israel", "Palestine", "Turkey"
       
      ) ~ "Asia",
  
      # --- Europe (includes Russia) ---
      cit_country %in% c(
        "United Kingdom", "Ireland", "France", "Germany", "Spain", "Portugal",
        "Italy", "Belgium", "Netherlands", "Luxembourg",
        "Switzerland", "Austria", "Poland", "Czech Republic",
        "Slovakia", "Hungary", "Slovenia", "Croatia",
        "Serbia", "Bosnia and Herzegovina", "Montenegro",
        "North Macedonia", "Albania", "Greece",
        "Bulgaria", "Romania", "Moldova", "Ukraine", "Belarus",
        "Lithuania", "Latvia", "Estonia",
        "Norway", "Sweden", "Finland", "Denmark", "Iceland",
        "Georgia", "Armenia", "Azerbaijan", "Russia"
      ) ~ "Europe",
  
      # --- Oceania ---
      cit_country %in% c(
        "Australia", "New Zealand", "Papua New Guinea",
        "Fiji", "Samoa", "Tonga", "Vanuatu", "Solomon Islands",
        "Kiribati", "Micronesia", "Palau",
        "Marshall Islands", "Nauru", "Tuvalu"
      ) ~ "Oceania",
  
      TRUE ~ NA_character_
    )
  )

```

```{r}
offices_clean <- offices |>
  drop_na(criminality, gender, age, region)

nrow(offices_clean)
```

## Logistic Regression (no spatial element)
```{r}
log_reg <- glm(deported ~ criminality + gender + age + region, 
               family = binomial,
               data = offices_clean)
```

```{r}
summary(log_reg)
```

Get residuals to map if I want to later
```{r}
offices_clean$log_resids <- log_reg$residuals
```

## Moran's I to see if the residuals are clustered
Based on our small p-value, we have enough evidence to reject the null hypothesis and conclude that there is spatial autocorrelation in the residuals. Although heteroskedasticity does not violate logistic regression assumptions, I am curious to see how the effects of various factors (e.g. criminality, region of citizenship) vary across field offices. 
```{r}
offices_grouped <- offices_clean |>
  group_by(office) |>
  summarize(avg_log_resid = mean(log_resids)) 


#coords <- st_coordinates(st_centroid(offices_grouped))

dist_matrix <- st_distance(offices_grouped)
w1 <- 1/dist_matrix

w1[is.infinite(w1)] <- 0
attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_log_resid, mat2listw(w1, style="W"))

```

## Hotspot Analysis on number of encounters
```{r}

```



## Clustering to use in Hierarchical Modeling
```{r}
coords <- st_coordinates(offices_clean)

set.seed(42069)
km <- kmeans(coords, centers = 100)

# 3. Add cluster ID to your data
offices_clean$cluster <- as.factor(km$cluster)
```




## Spatial Hierarchal Modeling
Hierarchical model with random intercepts and a random slope for criminality for each office and or
```{r}
# sample from the dataset bc it takes too long to run
set.seed(42069)
rand_sample <- offices_clean[sample(nrow(offices_clean), size = 5000), ]

glmer_reg <- glmer(deported ~ criminality + gender + scale(age) + region + (1 + criminality | aor / office), 
                   family = binomial,
                   data = rand_sample)
```

```{r}
rand_sample_df <- rand_sample |>
  st_drop_geometry() |>
  group_by(office)|>
  count()

rand_sample_df
```

```{r}
offices_clean |>
  st_drop_geometry() |>
  group_by(office)|>
  count() |>
  arrange(desc(n))

```



```{r}
summary(glmer_reg)
```

Random intercepts and slopes
```{r}
rand_effects_office <- ranef(glmer_reg)$office
rand_effects_aor <- ranef(glmer_reg$`aor:office`)

```

Map the office-level slopes
```{r}
rand_sample <- rand_sample |>
  group_by(office) |>
  left_join(rand_effects, by = c("office" = rownames(rand_sample)))

# convert to sf
rand_sample_sf <- st_as_sf(rand_sample, crs="ESRI:102010")
```

