---
title: "Modeling Framework"
format: html
editor: source
---
```{r}
library(tidyverse)
library(sf)
library(VGAM)
library(lubridate)
library(lme4)
library(spdep)
library(dbscan)
library(car)
library(corrplot)
library(GWmodel)
library(raster)
library(paletteer)
library(caret)
library(pROC)
#library(glmmTMB)
library(viridis)
library(emmeans)
select <- dplyr::select
```

Read in clustered data
```{r}
cnty_clus_offices <- st_read("data/cnty_clus_offices.gpkg")

# some preprocessing
cnty_clus_offices <- cnty_clus_offices |>
  drop_na(criminality, gender, age, region, avg_p_blk, avg_p_asn , avg_p_his, avg_p_nc, avg_pop_sqm)

cnty_clus_offices$deported <- factor(cnty_clus_offices$deported, levels = c(0,1))

cnty_clus_offices$region <- factor(cnty_clus_offices$region)

cnty_clus_offices$cluster <- factor(cnty_clus_offices$cluster)

cnty_clus_offices$gender <- factor(cnty_clus_offices$gender)

cnty_clus_offices$criminality <- factor(cnty_clus_offices$criminality)

cnty_clus_offices$criminality <- relevel(cnty_clus_offices$criminality, ref = "3 Other Immigration Violator")

cnty_clus_offices$cluster <- relevel(cnty_clus_offices$cluster, ref = "43")

# from https://www.americanimmigrationcouncil.org/wp-content/uploads/2025/01/cap_fact_sheet_8-1_fin_0.pdf 
cnty_clus_offices <- cnty_clus_offices |>
  mutate(program_type = factor(case_when(final_program == "ERO Criminal Alien Program" | 
                                           final_program == "Violent Criminal Alien Section" |
                                           final_program == "Law Enforcement Area Response Units" ~ "Criminal Alien Program (CAP)",
                                     final_program == "287G Program" | final_program == "287g Task Force" ~ "287(g)",
                                     TRUE ~  "Other")))
cnty_clus_offices$program_type <- relevel(cnty_clus_offices$program_type, ref = "Other")

latin_am_regions = c("Central America", "Caribbean", "Mexico", "South America")

cnty_clus_offices <- cnty_clus_offices |>
  mutate(latin_cit = case_when(region == "Mexico" ~ 1,
                                  region == "South America" ~ 1,
                                  region == "Central America" | region == "Caribbean" ~ 1,
                                  !(region %in% latin_am_regions) ~ 0)) 
```


### Check for multicollinearity
```{r}
numeric_vars <- cnty_clus_offices |>
  group_by(office) |>
  select(c(avg_p_blk, avg_p_asn , avg_p_his, avg_p_nc, avg_pop_sqm)) |>
  drop_na() |>
  st_drop_geometry()
```

None of correlations are above 0.8, so I can keep all variables in the model
```{r}
corr_mat <- cor(numeric_vars)
corrplot(corr_mat, type="upper", method="ellipse")
```
```{r}
corr_mat
```

## Basic logistic regression
```{r}
freq_table <- table(cnty_clus_offices$prog_287g)
prop.table(freq_table)
```

```{r}
#cnty_clus_offices$criminality <- factor(cnty_clus_offices$criminality)

log_reg <-glm(deported ~ criminality + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_p_nc + avg_pop_sqm + tps, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(log_reg , file="models/log_reg.rds")
```

```{r}
log_reg <- readRDS("models/log_reg.rds")
```

```{r}
summary(log_reg)
```
```{r}
log_reg2 <-glm(deported ~ criminality + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_pop_sqm + tps, 
               family = binomial,
               data = cnty_clus_offices)

summary(log_reg2)
```

```{r}
log_reg3 <-glm(deported ~ criminality + gender + age + latin_cit + avg_p_nc + avg_pop_sqm + tps, 
               family = binomial,
               data = cnty_clus_offices)

summary(log_reg3)
```

### Moran's I on basic logistic regression
```{r}
cnty_clus_offices$log_resids <- log_reg$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. We conclude that there is clustering behavior of the residuals. 
```{r}
offices_grouped <- cnty_clus_offices |>
  group_by(office) |>
  mutate(avg_log_resid = mean(log_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_log_resid, mat2listw(w1, style="W"))
```



```{r}
table(cnty_clus_offices$prog_287g)
```


## Interacting criminality with cluster 
Cluster reference group: 167
```{r}

cluster_reg_crim <- glm(deported ~ criminality + cluster + criminality:cluster + gender + age + latin_cit + tps, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_reg_crim, file="models/cluster_reg_crim.rds")
```


```{r}
cluster_reg_crim <- readRDS("models/cluster_reg_crim.rds")
```


```{r}
summary(cluster_reg_crim)
```

#### Get coefficients for the interaction term for convicted criminals


```{r}
coefs <- summary(cluster_reg_crim)$coefficients

# extract main effect of convicted criminal (same for all clusters)
cluster_fixed_crim <- coefs["criminality1 Convicted Criminal", "Estimate"]

# extract interaction terms (vary by cluster)
cluster_crim_int <- coefs[grep("^criminality1 Convicted Criminal:cluster", rownames(coefs)), , drop = FALSE]

# set non-significant interactions to 0 
cluster_crim_int[cluster_crim_int[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0


# create interaction dataframe
crim_df <- data.frame(
  cluster = sub("criminality1 Convicted Criminal:cluster", "", rownames(cluster_crim_int)),
  cluster_crim_int = cluster_crim_int[, "Estimate"]
)

# add the fixed effect (same for all clusters)
interaction_df <- interaction_df |>
  mutate(cluster_crim_fixed = cluster_fixed_crim)

# for clusters with no interaction term (reference cluster or not in model)
# add them with interaction = 0
cluster_office_df <- cnty_clus_offices

all_clusters <- unique(cluster_office_df$cluster)  # Get all cluster IDs from your data

interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(crim_df, by = "cluster") |>
  mutate(
    cluster_crim_fixed = cluster_fixed_crim,  
    cluster_crim_int = replace_na(cluster_crim_int, 0),
    cluster_crim_int_sig = ifelse(cluster_crim_int != 0, 1, 0)
  )

# calculate cluster-specific effect
interaction_df <- interaction_df |>
  mutate(cluster_crim_effect = cluster_crim_fixed + cluster_crim_int)

cluster_office_df <- cnty_clus_offices |>
  group_by(office) |>
  select(c(office, cluster, geom, aor)) |>
  slice(1) |>
  ungroup()

cluster_office_df <- cluster_office_df|>
  left_join(interaction_df, by = "cluster")
```


### Moran's I on cluster:criminality model
```{r}
cnty_clus_offices$clstr_crim_resids <- cluster_reg_crim$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- cnty_clus_offices |>
  group_by(office) |>
  mutate(avg_clstr_crim_resid = mean(clstr_crim_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_clstr_crim_resid, mat2listw(w1, style="W"))
```



## Interacting criminality with aor
s### exploration: calculate the proportion noncriminal per aor
```{r}
cnty_clus_offices |>
  group_by(aor)|>
  summarize(prop_noncrim = mean(criminality == "3 Other Immigration Violator")) |>
  arrange(desc(prop_noncrim))
```


```{r}

aor_reg <- glm(deported ~ criminality + aor + criminality:aor + gender + age + latin_cit + avg_p_blk + avg_p_asn  + avg_p_his + avg_p_nc + avg_pop_sqm + tps, 
               family = binomial,
               data = cnty_clus_offices)


saveRDS(aor_reg, file="models/aor_reg.rds")
```

```{r}
aor_reg <- readRDS("models/aor_reg.rds")
```


```{r}
summary(aor_reg)
```


#### Get coefficients for the interaction term for criminals at the AOR level
```{r}
coefs <- summary(aor_reg)$coefficients

# extract main effect of convicted criminal (same for all AORs)
aor_crim_int <- coefs["criminality1 Convicted Criminal", "Estimate"]

# extract interaction terms (vary by AOR)
aor_crim_int <- coefs[grep("^criminality1 Convicted Criminal:aor", rownames(coefs)), , drop = FALSE]

# set non-significant interactions to 0 
aor_crim_int[aor_crim_int[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0

# create interaction dataframe
aor_df <- data.frame(
  aor = sub("criminality1 Convicted Criminal:aor", "", rownames(aor_crim_int)),
  aor_crim_int = aor_crim_int[, "Estimate"]
)

# get all unique AORs from your data
all_aors <- unique(cluster_office_df$aor)

# add the fixed effect and handle missing interactions
aor_interaction_df <- data.frame(aor = all_aors) |>
  left_join(aor_df, by = "aor") |>
  mutate(
    aor_crim_fixed = aor_fixed_crim,  # same for all AORs
    aor_crim_int = ifelse(is.na(aor_crim_int), 0, aor_crim_int),
    aor_crim_int_sig = ifelse(aor_crim_int != 0, 1, 0)
  )

# calculate marginal effect
aor_interaction_df <- aor_interaction_df |>
  mutate(aor_crim_effect = aor_crim_fixed + aor_crim_int)


cluster_office_df <- cluster_office_df |>
  left_join(aor_interaction_df, by = "aor")
```

```{r}
aor_interaction_df_pct <- aor_interaction_df |>
  mutate(aor_crim_marg_pctdiff = exp(aor_crim_effect) * 100) 
```




### Moran's I on aor:criminality model
```{r}
cnty_clus_offices$aor_crim_resids <- aor_reg$residuals
```

At the alpha = 0.05 level, we fail to reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. (yay!)
```{r}
offices_grouped <- cnty_clus_offices |>
  group_by(office) |>
  mutate(avg_aor_crim_resid = mean(aor_crim_resids)) |>
  slice(1) |>
  ungroup()

office_coords <- st_coordinates(offices_grouped)

w1 <- 1/as.matrix(st_distance(offices_grouped ,
                              offices_grouped))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(offices_grouped$avg_aor_crim_resid, mat2listw(w1, style="W"))
```

## Interaction Age:cluster
```{r}
cluster_age <- glm(deported ~ criminality+ cluster +age:cluster+ gender + age + latin_cit + tps, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_age, file="models/cluster_age.rds")

```


```{r}
summary(cluster_age)
```

```{r}
cluster_age <- readRDS("models/cluster_age.rds")
```


#### Get coefficients for the interaction term for age
```{r}
coefs <- summary(cluster_age)$coefficients

# extract main effect of age
cluster_age_fixed <- coefs["age", "Estimate"]

# extract interaction terms
cluster_age_int <- coefs[grep(":age$", rownames(coefs)), , drop = FALSE]

# set non-significant interactions to 0
cluster_age_int[cluster_int_age[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0

# create interaction dataframe from model coefficients
age_df <- data.frame(
  cluster = sub("cluster", "", sub(":age$", "", rownames(cluster_int_age))),
  cluster_age_int = cluster_age_int[, "Estimate"]
)

# get all unique clusters
all_clusters <- unique(cluster_office_df$cluster)

# add fixed effect and handle missing interactions
age_interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(age_df, by = "cluster") |>
  mutate(
    cluster_age_fixed = cluster_age_fixed,
    cluster_age_int = replace_na(cluster_age_int, 0),
    cluster_age_sig = ifelse(cluster_age_int != 0 ,1, 0)
  )

# calculate cluster specific effect
age_interaction_df <- age_interaction_df |>
  mutate(cluster_age_effect = cluster_age_fixed + cluster_age_int)


# join to main dataframe
cluster_office_df <- cluster_office_df |>
  left_join(age_interaction_df, by = "cluster")
```



## Interaction LatinAmerica:cluster

```{r}
freq_table <- table(cnty_clus_offices$latin_cit)
prop.table(freq_table)
```

```{r}
cluster_latin_cit <- glm(deported ~ criminality + cluster +latin_cit:cluster+ gender + age + latin_cit + tps, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_latin_cit, file="models/cluster_latin_cit_bin.rds")
```

```{r}
summary(cluster_latin_cit)
```

```{r}
#cluster_latin_cit <- readRDS("models/cluster_latin_cit_bin.rds")
```

#### Get coefficients for Latin citizenship effects at cluster level
```{r}

coefs <- summary(cluster_latin_cit)$coefficients

# extract main effects for each Latin citizenship category
lat_fixed <- coefs["latin_cit", "Estimate"]

# extract interaction terms for each category
cluster_int_lat <- coefs[grep("latin_cit$", rownames(coefs)), , drop = FALSE]


# set non-significant interactions to 0 for each category
cluster_int_lat[cluster_int_lat[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0


# create interaction dataframes for each category
lat_df <- data.frame(
  cluster = sub("cluster", "", sub(":latin_cit$", "", rownames(cluster_int_lat))),
  cluster_int_lat = cluster_int_lat[, "Estimate"]
)


# get all unique clusters
all_clusters <- unique(cluster_office_df$cluster)

# combine all Latin citizenship effects
lat_interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(lat_df, by = "cluster") |>
  mutate(
    cluster_lat_fixed = lat_fixed,
    cluster_int_lat = ifelse(is.na(cluster_int_lat), 0, cluster_int_lat),
    cluster_lat_int_sig = ifelse(cluster_int_lat != 0, 1, 0)

  )

# calculate cluster-specific effects for each category
lat_interaction_df <- lat_interaction_df |>
  mutate(
    cluster_lat_effect = lat_fixed + cluster_int_lat,

  )

# join to main dataframe
cluster_office_df <- cluster_office_df |>
  left_join(lat_interaction_df, by = "cluster")
```

## Interaction tps:cluster

```{r}
cluster_tps <- glm(deported ~ criminality + cluster +tps:cluster+ gender + age + latin_cit + tps, 
               family = binomial,
               data = cnty_clus_offices)

saveRDS(cluster_tps, file="models/cluster_tps.rds")
```

```{r}
cluster_tps <- readRDS("models/cluster_tps.rds")
```

#### Get coefficients for TPS at cluster level
```{r}

coefs <- summary(cluster_tps)$coefficients

# extract main effects for each Latin citizenship category
cluster_tps_fixed <- coefs["tps", "Estimate"]

# extract interaction terms for each category
cluster_int_tps<- coefs[grep("tps$", rownames(coefs)), , drop = FALSE]


# set non-significant interactions to 0 for each category
cluster_int_tps[cluster_int_tps[, "Pr(>|z|)"] > 0.05, "Estimate"] <- 0


# create interaction dataframes for each category
tps_df <- data.frame(
  cluster = sub("cluster", "", sub(":tps$", "", rownames(cluster_int_tps))),
  cluster_int_tps = cluster_int_tps[, "Estimate"]
)


# get all unique clusters
all_clusters <- unique(cluster_office_df$cluster)

# combine all Latin citizenship effects
tps_interaction_df <- data.frame(cluster = all_clusters) |>
  left_join(tps_df, by = "cluster") |>
  mutate(
    # fixed effects (same for all clusters)
    cluster_tps_fixed = cluster_tps_fixed,
    cluster_int_tps = ifelse(is.na(cluster_int_tps), 0, cluster_int_tps),
    cluster_tps_int_sig = ifelse(cluster_int_tps != 0, 1, 0)

  )

# calculate cluster-specific effects for each category
tps_interaction_df <- tps_interaction_df |>
  mutate(
    cluster_tps_effect = cluster_tps_fixed + cluster_int_tps,

  )

# join to main dataframe
cluster_office_df <- cluster_office_df |>
  left_join(tps_interaction_df, by = "cluster")
```


### Transform interaction terms into percent change


```{r}
cluster_office_df_transf <- cluster_office_df |>
  mutate(across(
    .cols = c(cluster_crim_effect, aor_crim_effect, cluster_lat_effect, cluster_tps_effect, cluster_age_effect,
              cluster_crim_fixed, aor_crim_fixed, cluster_age_fixed, cluster_lat_fixed, cluster_tps_fixed),
    .fns = ~ (exp(.) - 1) * 100,
    .names = "{.col}_pctdiff"
  )) 

```



## Export coefficients

```{r}
saveRDS(cluster_office_df_transf, file = "mapping_coefs.rds") 
```


```{r}
st_write(cluster_office_df_transf, "data/mapping_coefs.gpkg")
```


### AIC model comparision
The model interacting cluster with criminality status has the lowest AIC
```{r}
print(AIC(cluster_reg_crim))
print(AIC(cluster_age))
print(AIC(cluster_latin_cit))
print(AIC(cluster_tps))
```

## Get predicted probabilities of deportation from cluster:criminality regression model
```{r}
pred_odds <- exp(predict(cluster_reg_crim))
pred_probs <- pred_odds / (pred_odds + 1)

cnty_clus_offices$pred_prob <- pred_probs
```

```{r}
cnty_clus_offices |>
  select(deported, pred_prob) |> 
  st_drop_geometry() |>
  slice_sample(n=10)
```


### Do train/test split to get confusion matrix, F1 score, and ROC curve
```{r}
set.seed(888)


# create train-test split, createDataPartition already shuffles your data
split_idx <- createDataPartition(cnty_clus_offices$deported, p = 0.8, list=FALSE)

encounters_train = cnty_clus_offices[split_idx, ]

encounters_test = cnty_clus_offices[-split_idx, ]



```

```{r}
cluster_reg_train <- glm(deported ~ criminality + cluster + criminality:cluster + gender + age + latin_cit + tps, 
               family = binomial,
               data = encounters_train)
```


```{r}
saveRDS(cluster_reg_train, file="models/cluster_reg_train.rds")
```

```{r}
#cluster_reg_train<- readRDS("cluster_reg_train.rds")
```


Get predictions from test set
```{r}
pred_probs <- predict(cluster_reg_train, newdata = encounters_test, type = "response")
```

ROC curve to decide where to place the "threshold" for classification
```{r}
roc_curve <- roc(response = encounters_test$deported, predictor = pred_probs)

plot(roc_curve, main = "ROC Curve", print.auc=TRUE,
     xlim = c(0, 1))
auc(roc_curve)
```


```{r}
# convert probabilities to predicted class
# TODO: ask about this
pred_dep <- ifelse(pred_probs >= 0.50, "1", "0")
pred_dep <- factor(pred_dep, levels= levels(encounters_test$deported))

# create confusion matrix
confusionMatrix(
  data = pred_dep,
  reference = encounters_test$deported,
  positive = "1" # well, not actually "positive"
)

```
```{r}
# proportion of deport predictions that were actually correct 
precision <- posPredValue(pred_dep, encounters_test$deported, positive = "1")

# proportion of correctly predicted deported cases
recall <- sensitivity(pred_dep, encounters_test$deported, positive = "1")

f1 <- (2*precision * recall) / (precision + recall)

print(precision)
print(recall)
print(f1)
```


```{r}
hist(pred_probs)
```


# GWR with aggregated offices
My computer crashes every time I try to compute the distance matrix because there are so many observations. So, my plan is to aggregate the data by office, but assign weights to offices with more datapoints. 

```{r}
# aggregate by offices
office_agg <- cnty_clus_offices |>
  group_by(office) |>
  reframe(prop_latin_cit = mean(latin_cit == 1),
         prop_crim_or_pending = mean(criminality == "1 Convicted Criminal" | criminality == "2 Pending Criminal Charges"),
         prop_female = mean(gender == "Female"),
         prop_male = mean(gender == "Male"),
         avg_prop_asn = mean(avg_p_asn),
         avg_prop_his = mean(avg_p_his),
         avg_prop_blk = mean(avg_p_blk),
         avg_prop_nc = mean(avg_p_nc),
         avg_pop_sqm = mean(avg_pop_sqm),
         avg_age = mean(age), 
         prop_deported = mean(deported == 1),
         num_encounters = num_encounters,
         office = office,
         aor = aor,
         geom = st_union(geom)) |>
   distinct(office, .keep_all = TRUE)

office_agg_sf <- st_as_sf(
  office_agg,
  sf_column_name = "geom",
  crs = st_crs(cnty_clus_offices)  
)
```



Note: summary statistics won't really make sense here because of the unequal distribution of encounters per office (weighting would need to be done first)

```{r}
office_agg |>
  summarize(min(num_encounters),
            max(num_encounters),
            mean(num_encounters),
            median(num_encounters),
            sd(num_encounters))
```


```{r}
office_agg_sp <- as(office_agg_sf, "Spatial")

bw <- bw.gwr(prop_deported ~ prop_latin_cit + prop_crim_or_pending + prop_female,
             data = office_agg_sp,
             approach = "CV",
             kernel = "bisquare",
             adaptive=TRUE)
```

## Fit OLS
```{r}
# ols first
ols_fit <- lm(prop_deported ~ prop_latin_cit + prop_crim_or_pending + prop_female + num_encounters + avg_prop_blk + avg_prop_asn  + avg_prop_his + avg_prop_nc + avg_pop_sqm,
             data = office_agg)

summary(ols_fit)
```

### Moran's I
```{r}
office_agg_sp$lm_resids <- ols_fit$residuals
```

At the alpha = 0.05 level, we reject the null hypothesis that the distribution of the average residuals across offices are spatially uncorrelated. 
```{r}
office_agg_sf <- st_as_sf(office_agg_sp)
office_coords <- st_coordinates(office_agg_sf)


w1 <- 1/as.matrix(st_distance(office_agg_sf,
                              office_agg_sf))
w1[is.infinite(w1)] <- 0

attr(w1, "class") <- NULL

moran.test(office_agg_sp$lm_resids, mat2listw(w1, style="W"))
```

## GWR
```{r}
# replaced population density with number of encounters
gwr_fit <- gwr.basic(prop_deported ~ prop_crim_or_pending + prop_latin_cit + prop_female + num_encounters,
             data = office_agg_sp,
             bw = bw,
             kernel = "bisquare",
             adaptive=TRUE)

gwr_results <- st_as_sf(gwr_fit$SDF)
gwr_fit 
```



```{r}
# alpha* adjustment

#effective number of parameters
pe <- 27.33118 
n <- nrow(office_agg_sp)
p <- 4

alpha_star <- 0.05/(1+pe - pe/(n*p))
alpha_star
df <- n - p - 1
t_star <- qt(1-alpha_star, df = df)
t_star 
```


```{r}
# assign significance
gwr_results <- gwr_results |>
  mutate(sig_crim = ifelse(abs(prop_crim_or_pending_TV) > t_star, prop_crim_or_pending, 0),
         sig_latin_cit = ifelse(abs(prop_latin_cit_TV) > t_star, prop_latin_cit, 0),
         #sig_his = ifelse(abs(avg_prop_his_TV) > t_star,avg_prop_his, 0),
         #sig_asn = ifelse(abs(avg_prop_asn_TV) > t_star, avg_prop_asn, 0),
         sig_female = ifelse(abs(prop_female_TV)> t_star, prop_female, 0))

hist(gwr_results$sig_crim)
hist(gwr_results$sig_latin_cit)
#hist(gwr_results$sig_his)
#hist(gwr_results$sig_asn)
hist(gwr_results$sig_female)

```

```{r}
# basic plot
gwr_results <- st_transform(gwr_results, st_crs(cnty_clus_offices))

# plot local coefficients as quintiles
# criminality
q_criminal <- quantile(gwr_results$sig_crim,
                       probs = seq(0, 1, 0.1),
                       na.rm=TRUE)

gwr_results$coef_crim_q <- case_when(
  gwr_results$sig_crim == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_crim,
                          breaks = unique(q_criminal),
                          include.lowest = TRUE))
)

plot_crim <- ggplot(gwr_results) +
  geom_sf(aes(color = coef_crim_q), linewidth = 0.2) +
  scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_crim, 
                                 breaks = unique(q_criminal), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size = 12) + 
  ggtitle("Criminal/Pending Criminal Charges Coef. (quintiles)")

plot_crim

```


```{r}
# 100% significant: space maters
# being from central america
# criminality
q_latin_cit <- quantile(gwr_results$sig_latin_cit,
                       probs = seq(0, 1, 0.2),
                       na.rm=TRUE)

gwr_results$coef_latin_cit_q <- case_when(
  gwr_results$sig_latin_cit == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_latin_cit,
                          breaks = unique(q_latin_cit),
                          include.lowest = TRUE))
)

plot_latin_cit <- ggplot(gwr_results) +
  geom_sf(aes(color = coef_latin_cit_q), linewidth = 0.2) +
  scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_latin_cit, 
                                 breaks = unique(q_latin_cit), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size = 12) + 
  ggtitle("Citizenship from a Latin American Country Coef. (quintiles)")

plot_latin_cit

```


```{r}
q_female <- quantile(gwr_results$sig_female,
                       probs = seq(0, 1, 0.2),
                       na.rm=TRUE)

gwr_results$coef_female_q <- case_when(
  gwr_results$sig_female == 0 ~ "Zero",
  TRUE ~ as.character(cut(gwr_results$sig_female,
                          breaks = unique(q_female),
                          include.lowest = TRUE))
)

plot_prop_female<- ggplot(gwr_results) +
  geom_sf(aes(color = coef_female_q), linewidth = 0.2) +
  scale_color_manual(
    values = c("Zero" = "gray50", 
               setNames(viridis(5, direction = -1), 
                       levels(cut(gwr_results$sig_female, 
                                 breaks = unique(q_female), 
                                 include.lowest = TRUE)))),
    name = "Coefficient"
  ) +
  theme_bw(base_size= 12) + 
  theme_bw() 
  ggtitle("Proportion Female Coef. (quintiles)")

plot_prop_female
```
